---
title: "KarensData"
author: "Josie Lesage"
date: "8/5/2020"
output: html_document
---


```{r setup, include=FALSE}
## community analysis
library(codyn)
library(MASS)
library(vegan)
library(lme4)

## misc/both/other
library(oz)
library(plotrix)
library(tidyverse)
library(gridExtra)
```

# File Goal

One of the key concerns about Karens' control plots is that they were released from grazing, and that this release may have affected the change in CTI over time. To potentially rebut this, I will examine the CTI of all of these plots over time, not just the control plots.

To begin, I'll re-clean all of the data, but will not isolate the control plot data. 


## Elkhorn
Data for grassland plots at Elkhorn Slough. Data were collected by Karen Holl and Grey Hayes between 1999 and 2018. For these analyses, I am using on the control plots (n=3), and removing all other treatments from the dataset. *Karen has suggested combining all of the Bromus species into one group*. 

Note that most of the code in the cleanup process was written before transitioning to R markdown and GitHub. 

```{r Importing Elkhorn, warning=FALSE, echo=FALSE}
elkspp <- read.csv("ComComp Data Cleanup/HollData/ElkhornSpecies.csv")
elkspp$species.code <- tolower(elkspp$species.code)

elk_guilds <- read.csv("ComComp Data Cleanup/HollData/ElkhornGuilds.csv")
elk99 <- read.csv("ComComp Data Cleanup/HollData/Elk_1999.csv",  skip = 4)
elk00 <- read.csv("ComComp Data Cleanup/HollData/Elk_2000.csv", skip = 3)
elk01 <- read.csv("ComComp Data Cleanup/HollData/Elk_2001.csv", skip = 3)
elk02 <- read.csv("ComComp Data Cleanup/HollData/Elk_2002.csv", skip = 3)
elk03 <- read.csv("ComComp Data Cleanup/HollData/Elk_2003.csv")
elk04 <- read.csv("ComComp Data Cleanup/HollData/Elk_2004.csv")
elk05 <- read.csv("ComComp Data Cleanup/HollData/Elk_2005.csv")
elk06 <- read.csv("ComComp Data Cleanup/HollData/Elk_2006.csv")
elk07 <- read.csv("ComComp Data Cleanup/HollData/Elk_2007.csv")
elk08 <- read.csv("ComComp Data Cleanup/HollData/Elk_2008.csv")
elk09 <- read.csv("ComComp Data Cleanup/HollData/Elk_2009.csv")
elk10 <- read.csv("ComComp Data Cleanup/HollData/Elk_2010.csv", skip = 1)
elk11 <- read.csv("ComComp Data Cleanup/HollData/Elk_2011.csv")
elk12 <- read.csv("ComComp Data Cleanup/HollData/Elk_2012.csv")
elk13 <- read.csv("ComComp Data Cleanup/HollData/Elk_2013.csv", skip = 2)
elk14 <- read.csv("ComComp Data Cleanup/HollData/Elk_2014.csv", skip = 1)
elk15 <- read.csv("ComComp Data Cleanup/HollData/Elk_2015.csv", skip = 1)
elk16 <- read.csv("ComComp Data Cleanup/HollData/Elk_2016.csv", skip = 1)
elk17 <- read.csv("ComComp Data Cleanup/HollData/Elk_2017.csv", skip = 1)
elk18 <- read.csv("ComComp Data Cleanup/HollData/Elk_2018.csv", skip = 1)

```

Because these data are in separate files (one year per file), I need to clean the data and compile all of the years into a single object. To clean the data, I selected only the Elkhorn plots (these files also contained the data from two other sites; UCSC and Swanton, removed the non-control plots, restructured the data to make sure only non-zero datapoints were included, and added the year.

I then combined the file with species guild information to each yearly object. 

```{r Cleaning Elkhorn, warning=FALSE, echo=FALSE}
elk99 <- elk99 %>%
  filter(site == "elk") %>%
  mutate(year = "1999") 
elk99 <- gather(elk99, "anaarv9":"bare9", key = "species.code", value = "intercepts")
elk99$species.code <- substr(elk99$species.code, 1, nchar(elk99$species.code)-1) 
elk99$species.code <- tolower(elk99$species.code)
elk99 <- left_join(elk99, elkspp, by = "species.code")
elk99$intercepts <- as.numeric(elk99$intercepts)

elk00 <- elk00 %>%
  filter(site == "elk") %>%
  select(site:naspul) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"naspul", key = "species.code", value = "intercepts") %>%
  mutate(year = "2000")
elk00$species.code <- tolower(elk00$species.code)
elk00 <- left_join(elk00, elkspp, by = "species.code")
elk00 <- spread(elk00, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
elk00$intercepts <- as.numeric(elk00$intercepts)

elk01 <- elk01 %>%
  filter(site == "elk") %>%
  select(site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2001")
elk01$species.code <- tolower(elk01$species.code)
elk01 <- left_join(elk01, elkspp, by = "species.code")
elk01 <- spread(elk01, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
elk01$intercepts <- as.numeric(elk01$intercepts)

elk02 <- elk02 %>%
  filter(Site == "elk") %>%
  select(Site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2002")
names(elk02) <- tolower(names(elk02))
elk02 <- elk02 %>%
  rename(site = site,
         rep = replicate,
         freq = frequency,
         treat = treatment) 
elk02$intercepts <- as.numeric(elk02$intercepts)
elk02 <- left_join(elk02, elkspp, by = "species.code")

elk03 <- elk03 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2003")
names(elk03) <- tolower(names(elk03))
elk03 <- left_join(elk03, elkspp, by = "species.code")

elk04 <- elk04 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2004") 
names(elk04) <- tolower(names(elk04))
elk04 <- left_join(elk04, elkspp, by = "species.code")

elk05 <- elk05 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2005")
names(elk05) <- tolower(names(elk05))
elk05 <- left_join(elk05, elkspp, by = "species.code")

elk06 <- elk06 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2006")
names(elk06) <- tolower(names(elk06))
elk06 <- left_join(elk06, elkspp, by = "species.code")

elk07 <- elk07 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2007")
names(elk07) <- tolower(names(elk07))
elk07 <- left_join(elk07, elkspp, by = "species.code")

elk08 <- elk08 %>%
  filter(SITE == "elk") %>%
  select(SITE:hormur) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"hormur", key = "species.code", value = "intercepts") %>%
  mutate(year = "2008")
names(elk08) <- tolower(names(elk08))
elk08 <- left_join(elk08, elkspp, by = "species.code")

elk09 <- elk09 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2009")
names(elk09) <- tolower(names(elk09))
elk09 <- left_join(elk09, elkspp, by = "species.code")

elk10 <- elk10 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2010")
names(elk10) <- tolower(names(elk10))
elk10$intercepts <- as.numeric(elk10$intercepts)
elk10 <- left_join(elk10, elkspp, by = "species.code")

elk11 <- elk11 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2011")
names(elk11) <- tolower(names(elk11))
elk11$intercepts <- as.numeric(elk11$intercepts)
elk11 <- left_join(elk11, elkspp, by = "species.code")

elk12 <- elk12 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2012")
names(elk12) <- tolower(names(elk12))
elk12$intercepts <- as.numeric(elk12$intercepts)
elk12 <- left_join(elk12, elkspp, by = "species.code")

elk13 <- elk13 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2013")
names(elk13) <- tolower(names(elk13))
elk13$intercepts <- as.numeric(elk13$intercepts)
elk13 <- left_join(elk13, elkspp, by = "species.code")

elk14 <- elk14 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2014")
names(elk14) <- tolower(names(elk14))
elk14 <- left_join(elk14, elkspp, by = "species.code")

elk15 <- elk15 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2015")
names(elk15) <- tolower(names(elk15))
elk15 <- left_join(elk15, elkspp, by = "species.code")

elk16 <- elk16 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpia) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpia", key = "species.code", value = "intercepts") %>%
  mutate(year = "2016")
names(elk16) <- tolower(names(elk16))
elk16 <- left_join(elk16, elkspp, by = "species.code")

elk17 <- elk17 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpia) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpia", key = "species.code", value = "intercepts") %>%
  mutate(year = "2017")
names(elk17) <- tolower(names(elk17))
elk17 <- left_join(elk17, elkspp, by = "species.code")


elk18 <- elk18 %>%
  filter(SITE == "elk") %>%
  select(SITE:vulpia) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpia", key = "species.code", value = "intercepts") %>%
  mutate(year = "2018")
names(elk18) <- tolower(names(elk18))
elk18 <- left_join(elk18, elkspp, by = "species.code")

```

At this point, all that is left to do is to combine the objects so that we have one large object with each year of data in it. 

```{r Compiling Elkhorn, warning=FALSE, echo=FALSE}
elk_99_00 <- full_join(elk99, elk00, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))

elk_99_01 <- full_join(elk_99_00, elk01, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))
remove(elk_99_00)

elk_99_02 <- full_join(elk_99_01, elk02, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_01)

elk_99_03 <- full_join(elk_99_02, elk03, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_02)

elk_99_04 <- full_join(elk_99_03, elk04, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_03)

elk_99_05 <- full_join(elk_99_04, elk05, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_04)

elk_99_06 <- full_join(elk_99_05, elk06, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_05)

elk_99_07 <- full_join(elk_99_06, elk07, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_06)

elk_99_08 <- full_join(elk_99_07, elk08, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_07)

elk_99_09 <- full_join(elk_99_08, elk09, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_08)

elk_99_10 <- full_join(elk_99_09, elk10, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_09)

elk_99_11 <- full_join(elk_99_10, elk11, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_10)

elk_99_12 <- full_join(elk_99_11, elk12, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_11)

elk_99_13 <- full_join(elk_99_12, elk13, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_12)

elk_99_14 <- full_join(elk_99_13, elk14, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_13)

elk_99_15 <- full_join(elk_99_14, elk15, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_14)

elk_99_16 <- full_join(elk_99_15, elk16, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_15)

elk_99_17 <- full_join(elk_99_16, elk17, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(elk_99_16)

elk_99_18 <- full_join(elk_99_17, elk18, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts")) 
remove(elk_99_17)
remove(elk99)
remove(elk00)
remove(elk01)
remove(elk02)
remove(elk03)
remove(elk04)
remove(elk05)
remove(elk06)
remove(elk07)
remove(elk08)
remove(elk09)
remove(elk10)
remove(elk11)
remove(elk12)
remove(elk13)
remove(elk14)
remove(elk15)
remove(elk16)
remove(elk17)
remove(elk18)
```

```{r Elkhorn graph prep, warning=FALSE, echo=FALSE}
elk_99_18_species <- elk_99_18 %>%
  filter(intercepts > 0) %>%
  select("freq", "treat", "rep", "year", "species.name", "intercepts", "guild") %>%
  mutate(site = "Elkhorn") 
elk_99_18_species$species.name <- as.character(elk_99_18_species$species.name)
elk_99_18_species$species.name <- as.factor(elk_99_18_species$species.name)
elk_99_18_species$year <- as.integer(elk_99_18_species$year)
elk_99_18_species$year <- as.numeric(elk_99_18_species$year)

elkhorn_data <- elk_99_18_species %>%
  mutate(species.name = fct_recode(species.name, 
                                   'Vulpia myuros' = 'Festuca myuros'))

unique(elkhorn_data_tidy$treat)

# fixing all of the treatment types to align with one another
elkhorn_data <- elkhorn_data %>%
  mutate(freq = fct_recode(freq,
                           "CL1" = "cl1",
                           "CL1" = "CL1",
                           "CL1" = "1x clip",
                           "CL2" = "CL2",
                           "CL2" = "cl2",
                           "CL2" = "2x clip",
                           "CL3" = "cl3",
                           "CL3" = "CL3",
                           "CL3" = "3x clip",
                           "Control" = "con",
                           "Control" = "CONTROL",
                           "Grazed" = "graz",
                           "Grazed" = "gra",
                           "Grazed" = "GRAZED"),
         treat = fct_recode(treat,
                           "LL" = "ll",
                           "RE" = "re",
                           "SD" = "sd"))

```

```{r Removing extra elkhorn objects}
remove(elk_99_18_guilds)
remove(elk_99_18_species)
remove(elk_guilds)
remove(elkspp)
remove(elk_99_18)
```

```{r Importing Swanton, warning=FALSE, echo=FALSE}
Swanspp <- read.csv("ComComp Data Cleanup/HollData/ElkhornSpecies.csv")
Swanspp$species.code <- tolower(Swanspp$species.code)
Swan_guilds <- read.csv("ComComp Data Cleanup/HollData/ElkhornGuilds.csv")

Swan99 <- read.csv("ComComp Data Cleanup/HollData/Elk_1999.csv",  skip = 4)
Swan00 <- read.csv("ComComp Data Cleanup/HollData/Elk_2000.csv", skip = 3)
Swan01 <- read.csv("ComComp Data Cleanup/HollData/Elk_2001.csv", skip = 3)
Swan02 <- read.csv("ComComp Data Cleanup/HollData/Elk_2002.csv", skip = 3)
Swan03 <- read.csv("ComComp Data Cleanup/HollData/Elk_2003.csv")
Swan04 <- read.csv("ComComp Data Cleanup/HollData/Elk_2004.csv")
Swan05 <- read.csv("ComComp Data Cleanup/HollData/Elk_2005.csv")
Swan06 <- read.csv("ComComp Data Cleanup/HollData/Elk_2006.csv")
Swan07 <- read.csv("ComComp Data Cleanup/HollData/Elk_2007.csv")
Swan08 <- read.csv("ComComp Data Cleanup/HollData/swanton_08.csv", skip = 1)
Swan09 <- read.csv("ComComp Data Cleanup/HollData/Elk_2009.csv")
Swan10 <- read.csv("ComComp Data Cleanup/HollData/Elk_2010.csv", skip = 1)
Swan11 <- read.csv("ComComp Data Cleanup/HollData/Elk_2011.csv")
Swan12 <- read.csv("ComComp Data Cleanup/HollData/Elk_2012.csv")

```

```{r Cleaning Swanton, warning=FALSE, echo=FALSE}
Swan99 <- Swan99 %>%
  filter(site == "swan") %>%
  mutate(year = "1999")
Swan99 <- gather(Swan99, "anaarv9":"bare9", key = "species.code", value = "intercepts")
Swan99$species.code <- substr(Swan99$species.code, 1, nchar(Swan99$species.code)-1) 
Swan99$species.code <- tolower(Swan99$species.code)
Swan99 <- left_join(Swan99, Swanspp, by = "species.code")
Swan99$intercepts <- as.numeric(Swan99$intercepts)

Swan00 <- Swan00 %>%
  filter(site == "swa") %>%
  select(site:naspul) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"naspul", key = "species.code", value = "intercepts") %>%
  mutate(year = "2000")
Swan00$species.code <- tolower(Swan00$species.code)
Swan00 <- left_join(Swan00, Swanspp, by = "species.code")
Swan00 <- spread(Swan00, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
Swan00$intercepts <- as.numeric(Swan00$intercepts)

Swan01 <- Swan01 %>%
  filter(site == "swan") %>%
  select(site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2001")
Swan01$species.code <- tolower(Swan01$species.code)
Swan01 <- left_join(Swan01, Swanspp, by = "species.code")
Swan01 <- spread(Swan01, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
Swan01$intercepts <- as.numeric(Swan01$intercepts)

Swan02 <- Swan02 %>%
  filter(Site == "swa") %>%
  select(Site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2002")
names(Swan02) <- tolower(names(Swan02))
Swan02 <- Swan02 %>%
  rename(site = site,
         freq = frequency,
         treat = treatment,
         rep = replicate) 
Swan02$intercepts <- as.numeric(Swan02$intercepts)
Swan02 <- left_join(Swan02, Swanspp, by = "species.code")

Swan03 <- Swan03 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2003")
names(Swan03) <- tolower(names(Swan03))
Swan03 <- left_join(Swan03, Swanspp, by = "species.code")

Swan04 <- Swan04 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2004") 
names(Swan04) <- tolower(names(Swan04))
Swan04 <- left_join(Swan04, Swanspp, by = "species.code")

Swan05 <- Swan05 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2005")
names(Swan05) <- tolower(names(Swan05))
Swan05 <- left_join(Swan05, Swanspp, by = "species.code")

Swan06 <- Swan06 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2006")
names(Swan06) <- tolower(names(Swan06))
Swan06 <- left_join(Swan06, Swanspp, by = "species.code")

Swan07 <- Swan07 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2007")
names(Swan07) <- tolower(names(Swan07))
Swan07 <- left_join(Swan07, Swanspp, by = "species.code")

Swan08 <- Swan08 %>%
  filter(SITE == "SWANTON") %>%
  select(SITE:spearv) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"spearv", key = "species.code", value = "intercepts") %>%
  mutate(year = "2008")
names(Swan08) <- tolower(names(Swan08)) 
Swan08 <- left_join(Swan08, Swanspp, by = "species.code")

Swan09 <- Swan09 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2009")
names(Swan09) <- tolower(names(Swan09))
Swan09 <- left_join(Swan09, Swanspp, by = "species.code")

Swan10 <- Swan10 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2010")
names(Swan10) <- tolower(names(Swan10))
Swan10$intercepts <- as.numeric(Swan10$intercepts)
Swan10 <- left_join(Swan10, Swanspp, by = "species.code")

Swan11 <- Swan11 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2011")
names(Swan11) <- tolower(names(Swan11))
Swan11$intercepts <- as.numeric(Swan11$intercepts)
Swan11 <- left_join(Swan11, Swanspp, by = "species.code")

Swan12 <- Swan12 %>%
  filter(SITE == "swa") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2012")
names(Swan12) <- tolower(names(Swan12))
Swan12$intercepts <- as.numeric(Swan12$intercepts)
Swan12 <- left_join(Swan12, Swanspp, by = "species.code")
```

Once the data is cleaned up on a yearly basis, we can combine the years into a single dataframe. 

```{r Compiling Swanton, warning=FALSE, echo=FALSE}
Swan_99_00 <- full_join(Swan99, Swan00, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))

Swan_99_01 <- full_join(Swan_99_00, Swan01, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))
remove(Swan_99_00)

Swan_99_02 <- full_join(Swan_99_01, Swan02, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_01)

Swan_99_03 <- full_join(Swan_99_02, Swan03, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_02)

Swan_99_04 <- full_join(Swan_99_03, Swan04, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_03)

Swan_99_05 <- full_join(Swan_99_04, Swan05, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_04)

Swan_99_06 <- full_join(Swan_99_05, Swan06, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_05)

Swan_99_07 <- full_join(Swan_99_06, Swan07, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_06)

Swan_99_08 <- full_join(Swan_99_07, Swan08, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_07)

Swan_99_09 <- full_join(Swan_99_08, Swan09, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_08)

Swan_99_10 <- full_join(Swan_99_09, Swan10, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_09)

Swan_99_11 <- full_join(Swan_99_10, Swan11, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_10)

Swan_99_12 <- full_join(Swan_99_11, Swan12, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(Swan_99_11)

remove(Swan99)
remove(Swan00)
remove(Swan01)
remove(Swan02)
remove(Swan03)
remove(Swan04)
remove(Swan05)
remove(Swan06)
remove(Swan07)
remove(Swan08)
remove(Swan09)
remove(Swan10)
remove(Swan11)
remove(Swan12)
```

```{r Swanton Graph Prep, warning=FALSE, echo=FALSE}
Swan_99_12_species <- Swan_99_12 %>%
  filter(intercepts > 0) %>%
  select("freq", "treat", "rep", "year", "species.name", "intercepts", "guild") %>%
  mutate(site = "Swan")
Swan_99_12_species$species.name <- as.character(Swan_99_12_species$species.name)
Swan_99_12_species$species.name <- as.factor(Swan_99_12_species$species.name)
Swan_99_12_species$year <- as.integer(Swan_99_12_species$year)
Swan_99_12_species$year <- as.numeric(Swan_99_12_species$year)

swanton_data <- Swan_99_12_species %>%
  mutate(species.name = fct_recode(species.name, 
                                   'Vulpia myuros' = 'Festuca myuros'))

# fixing all of the treatment types to align with one another
swanton_data <- swanton_data %>%
  mutate(freq = fct_recode(freq,
                           "CL1" = "cl1",
                           "CL1" = "CL1",
                           "CL1" = "1x clip",
                           "CL2" = "CL2",
                           "CL2" = "cl2",
                           "CL2" = "2x clip",
                           "CL3" = "cl3",
                           "CL3" = "CL3",
                           "CL3" = "3x clip",
                           "Control" = "con",
                           "Control" = "CONTROL",
                           "Control" = "CON",
                           "Grazed" = "graz",
                           "Grazed" = "gra",
                           "Grazed" = "GRA",
                           "Grazed" = "GRAZED"),
         treat = fct_recode(treat,
                           "LL" = "ll",
                           "RE" = "re",
                           "SD" = "sd"))

```

```{r Importing UCSC, warning=FALSE, echo=FALSE}

UCSCspp <- read.csv("ComComp Data Cleanup/HollData/ElkhornSpecies.csv")
UCSCspp$species.code <- tolower(UCSCspp$species.code)
UCSC_guilds <- read.csv("ComComp Data Cleanup/HollData/ElkhornGuilds.csv")

UCSC99 <- read.csv("ComComp Data Cleanup/HollData/Elk_1999.csv",  skip = 4)
UCSC00 <- read.csv("ComComp Data Cleanup/HollData/Elk_2000.csv", skip = 3)
UCSC01 <- read.csv("ComComp Data Cleanup/HollData/Elk_2001.csv", skip = 3)
UCSC02 <- read.csv("ComComp Data Cleanup/HollData/Elk_2002.csv", skip = 3)
UCSC03 <- read.csv("ComComp Data Cleanup/HollData/Elk_2003.csv")
UCSC04 <- read.csv("ComComp Data Cleanup/HollData/Elk_2004.csv")
UCSC05 <- read.csv("ComComp Data Cleanup/HollData/Elk_2005.csv")
UCSC06 <- read.csv("ComComp Data Cleanup/HollData/Elk_2006.csv")
UCSC07 <- read.csv("ComComp Data Cleanup/HollData/Elk_2007.csv")
UCSC08 <- read.csv("ComComp Data Cleanup/HollData/ucsc_08.csv")
UCSC09 <- read.csv("ComComp Data Cleanup/HollData/Elk_2009.csv")
UCSC10 <- read.csv("ComComp Data Cleanup/HollData/Elk_2010.csv", skip = 1)
UCSC11 <- read.csv("ComComp Data Cleanup/HollData/Elk_2011.csv")
UCSC12 <- read.csv("ComComp Data Cleanup/HollData/Elk_2012.csv")
```

```{r Cleaning UCSC, warning=FALSE, echo=FALSE}
UCSC99 <- UCSC99 %>%
  filter(site == "ucsc") %>%
  mutate(year = "1999")
UCSC99 <- gather(UCSC99, "anaarv9":"bare9", key = "species.code", value = "intercepts")
UCSC99$species.code <- substr(UCSC99$species.code, 1, nchar(UCSC99$species.code)-1) 
UCSC99$species.code <- tolower(UCSC99$species.code)
UCSC99 <- left_join(UCSC99, UCSCspp, by = "species.code")
UCSC99$intercepts <- as.numeric(UCSC99$intercepts)

UCSC00 <- UCSC00 %>%
  filter(site == "ucsc") %>%
  select(site:naspul) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"naspul", key = "species.code", value = "intercepts") %>%
  mutate(year = "2000")
UCSC00$species.code <- tolower(UCSC00$species.code)
UCSC00 <- left_join(UCSC00, UCSCspp, by = "species.code")
UCSC00 <- spread(UCSC00, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
UCSC00$intercepts <- as.numeric(UCSC00$intercepts)

UCSC01 <- UCSC01 %>%
  filter(site == "ucsc") %>%
  select(site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2001")
UCSC01$species.code <- tolower(UCSC01$species.code)
UCSC01 <- left_join(UCSC01, UCSCspp, by = "species.code")
UCSC01 <- spread(UCSC01, key = plot, value = intercepts) %>%
  rename(plot1 = "1", 
         plot2 = "2",
         plot3 = "3",
         plot4 = "4") %>%
  mutate(intercepts = (plot1 + plot2 + plot3 + plot4)) %>%
  select(-plot1, -plot2, -plot3, -plot4)
UCSC01$intercepts <- as.numeric(UCSC01$intercepts)

UCSC02 <- UCSC02 %>%
  filter(Site == "ucsc") %>%
  select(Site:thatch) %>%
  replace(is.na(.), 0) %>%
  gather("anaarv":"thatch", key = "species.code", value = "intercepts") %>%
  mutate(year = "2002")
names(UCSC02) <- tolower(names(UCSC02))
UCSC02 <- UCSC02 %>%
  rename(site = site,
         freq = frequency,
         treat = treatment,
         rep = replicate) 
UCSC02$intercepts <- as.numeric(UCSC02$intercepts)
UCSC02 <- left_join(UCSC02, UCSCspp, by = "species.code")

UCSC03 <- UCSC03 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2003")
names(UCSC03) <- tolower(names(UCSC03))
UCSC03 <- left_join(UCSC03, UCSCspp, by = "species.code")

UCSC04 <- UCSC04 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2004") 
names(UCSC04) <- tolower(names(UCSC04))
UCSC04 <- left_join(UCSC04, UCSCspp, by = "species.code")

UCSC05 <- UCSC05 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2005")
names(UCSC05) <- tolower(names(UCSC05))
UCSC05 <- left_join(UCSC05, UCSCspp, by = "species.code")

UCSC06 <- UCSC06 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2006")
names(UCSC06) <- tolower(names(UCSC06))
UCSC06 <- left_join(UCSC06, UCSCspp, by = "species.code")

UCSC07 <- UCSC07 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2007")
names(UCSC07) <- tolower(names(UCSC07))
UCSC07 <- left_join(UCSC07, UCSCspp, by = "species.code")

UCSC08 <- UCSC08 %>%
  filter(SITE == "UCSC") %>%
  select(SITE:brohor) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"brohor", key = "species.code", value = "intercepts") %>%
  mutate(year = "2008")
names(UCSC08) <- tolower(names(UCSC08))
UCSC08 <- left_join(UCSC08, UCSCspp, by = "species.code")

UCSC09 <- UCSC09 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2009")
names(UCSC09) <- tolower(names(UCSC09))
UCSC09 <- left_join(UCSC09, UCSCspp, by = "species.code")

UCSC10 <- UCSC10 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2010")
names(UCSC10) <- tolower(names(UCSC10))
UCSC10$intercepts <- as.numeric(UCSC10$intercepts)
UCSC10 <- left_join(UCSC10, UCSCspp, by = "species.code")

UCSC11 <- UCSC11 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2011")
names(UCSC11) <- tolower(names(UCSC11))
UCSC11$intercepts <- as.numeric(UCSC11$intercepts)
UCSC11 <- left_join(UCSC11, UCSCspp, by = "species.code")

UCSC12 <- UCSC12 %>%
  filter(SITE == "ucsc") %>%
  select(SITE:vulpsp, -Comments) %>%
  replace(is.na(.), 0) %>%
  gather("aircar":"vulpsp", key = "species.code", value = "intercepts") %>%
  mutate(year = "2012")
names(UCSC12) <- tolower(names(UCSC12))
UCSC12$intercepts <- as.numeric(UCSC12$intercepts)
UCSC12 <- left_join(UCSC12, UCSCspp, by = "species.code")
```

```{r Compiling UCSC, warning=FALSE, echo=FALSE}
UCSC_99_00 <- full_join(UCSC99, UCSC00, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))

UCSC_99_01 <- full_join(UCSC_99_00, UCSC01, by = c("site", "freq", "treat", "rep", "year",
                                          "species.code", "species.name", "guild",
                                          "nat.exo", "ann.perr", "lifeform",
                                          "intercepts"))
remove(UCSC_99_00)

UCSC_99_02 <- full_join(UCSC_99_01, UCSC02, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_01)

UCSC_99_03 <- full_join(UCSC_99_02, UCSC03, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_02)

UCSC_99_04 <- full_join(UCSC_99_03, UCSC04, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_03)

UCSC_99_05 <- full_join(UCSC_99_04, UCSC05, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_04)

UCSC_99_06 <- full_join(UCSC_99_05, UCSC06, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_05)

UCSC_99_07 <- full_join(UCSC_99_06, UCSC07, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_06)

UCSC_99_08 <- full_join(UCSC_99_07, UCSC08, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_07)

UCSC_99_09 <- full_join(UCSC_99_08, UCSC09, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_08)

UCSC_99_10 <- full_join(UCSC_99_09, UCSC10, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_09)

UCSC_99_11 <- full_join(UCSC_99_10, UCSC11, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_10)

UCSC_99_12 <- full_join(UCSC_99_11, UCSC12, by = c("site", "freq", "treat", "rep", "year",
                                                "species.code", "species.name", "guild",
                                                "nat.exo", "ann.perr", "lifeform",
                                                "intercepts"))
remove(UCSC_99_11)

remove(UCSC99)
remove(UCSC00)
remove(UCSC01)
remove(UCSC02)
remove(UCSC03)
remove(UCSC04)
remove(UCSC05)
remove(UCSC06)
remove(UCSC07)
remove(UCSC08)
remove(UCSC09)
remove(UCSC10)
remove(UCSC11)
remove(UCSC12)
```

```{r UCSC Graph Prep, warning=FALSE, echo=FALSE}
UCSC_99_12_species <- UCSC_99_12 %>%
  filter(intercepts > 0) %>%
  select("freq", "treat", "rep", "year", "species.name", "intercepts", "guild") %>%
  mutate(site = "UCSC")
UCSC_99_12_species$species.name <- as.character(UCSC_99_12_species$species.name)
UCSC_99_12_species$species.name <- as.factor(UCSC_99_12_species$species.name)
UCSC_99_12_species$year <- as.integer(UCSC_99_12_species$year)
UCSC_99_12_species$year <- as.numeric(UCSC_99_12_species$year)

ucsc_data <- UCSC_99_12_species %>%
    mutate(species.name = fct_recode(species.name, 
                                   'Vulpia myuros' = 'Festuca myuros'))

ucsc_data <- ucsc_data %>%
  mutate(freq = fct_recode(freq,
                           "CL1" = "cl1",
                           "CL1" = "CL1",
                           "CL1" = "1x clip",
                           "CL2" = "CL2",
                           "CL2" = "cl2",
                           "CL2" = "2x clip",
                           "CL3" = "cl3",
                           "CL3" = "CL3",
                           "CL3" = "3x clip",
                           "Control" = "con",
                           "Control" = "CONTROL",
                           "Control" = "CON",
                           "Grazed" = "graz",
                           "Grazed" = "gra",
                           "Grazed" = "GRA",
                           "Grazed" = "GRAZED",
                           "Grazed" = "grazed"),
         treat = fct_recode(treat,
                           "LL" = "ll",
                           "RE" = "re",
                           "SD" = "sd"))

```



```{r Removing extra objects}
remove(UCSC_99_12)
remove(UCSC_99_12_species)
remove(UCSCspp)
remove(UCSC_guilds)
remove(Swan_99_12)
remove(Swan_99_12_species)
remove(Swanspp)
remove(Swan_guilds)
```


```{r saving data files}
# write.csv(elkhorn_data, "Analysis/ComData/elkhorn_data_allplots.csv")
# write.csv(swanton_data, "Analysis/ComData/swanton_data_allplots.csv")
# write.csv(ucsc_data, "Analysis/ComData/ucsc_data_allplots.csv")
```

# Combine all of the sites into a single dataset

```{r combine data}

## elkhorn
raw_elk <- elkhorn_data %>%
  filter(species.name != "Moss",
         species.name != "Thatch",
         species.name != "Bare ground",
         species.name != "Dung") %>%
  mutate(site = "Elkhorn")
rel_elk <- raw_elk %>%
  select(freq, treat, site, year, rep, species.name, intercepts, guild) %>%
  group_by(freq, treat, rep, year) %>%
  mutate(sumcov = sum(intercepts),
         relcov = intercepts/sumcov,
         datacol = "pt_intercept") %>%
  rename(plot = rep,
         cover = intercepts)
rel_elk$plot <- as.factor(rel_elk$plot)

## swanton
raw_swan <- swanton_data %>%
  filter(species.name != "Moss",
         species.name != "Thatch",
         species.name != "Bare ground",
         species.name != "Dung") %>%
  mutate(site = "Swanton")
rel_swan <- raw_swan %>%
  select(freq, treat, site, year, rep, species.name, intercepts, guild) %>%
  group_by(freq, treat, rep, year) %>%
  mutate(sumcov = sum(intercepts),
         relcov = intercepts/sumcov,
         datacol = "pt_intercept") %>%
  rename(cover = intercepts,
         plot = rep)
rel_swan$plot <- as.factor(rel_swan$plot)

## ucsc
raw_ucsc <- ucsc_data %>%
  filter(species.name != "Moss",
         species.name != "Thatch",
         species.name != "Bare ground",
         species.name != "Dung") %>%
  mutate(site = "UCSC")
rel_ucsc <- raw_ucsc %>%
  select(freq, treat, site, year, rep, species.name, intercepts, guild) %>%
  group_by(freq, treat, rep, year) %>%
  mutate(sumcov = sum(intercepts),
         relcov = intercepts/sumcov,
         datacol = "pt_intercept") %>%
  rename(cover = intercepts,
         plot = rep)
rel_ucsc$plot <- as.factor(rel_ucsc$plot)

# combine
rel_all <- full_join(rel_elk, rel_swan)
rel_all <- full_join(rel_all, rel_ucsc)

rel_all <- rel_all %>%
  mutate(species.name.2 = str_squish(species.name)) %>%
  select(-species.name) %>% rename(species.name = species.name.2)


# write.csv(rel_all, "Analysis/ComData/AllKarenData_Relative.csv")
```

## Calculating CTI

```{r load CTI packages}
## CTI analysis
library(CoordinateCleaner)
library(countrycode)
library(mapr)
library(maptools)
library(raster)
library(rgbif)
library(scrubr)
library(sp)
library(spocc)
require(rgbif)
```



```{r getting CTI data}
all_species <- unique(rel_all$species.name)
all_species <- as.data.frame(all_species)
all_species$all_species <- paste0("'", all_species$all_species, "',")
cat(format_delim(all_species, delim = ""))

```


```{r}
all_species_list <- c('Erodium botrys',
                      'Erodium moschatum',
                      'Geranium dissectum',
                      'Stellaria media',
                      'Trifolium angustifolium',
                      'Trifolium dubium',
                      'Trifolium subterraneum',
                      'Hypochaeris radicata',
                      'Plantago lanceolata',
                      'Rumex acetosella',
                      'Madia sativa',
                      'Taraxia ovata',
                      'Aira caryophyllea',
                      'Bromus diandrus',
                      'Bromus hordeaceus',
                      'Festuca perennis',
                      'Vulpia myuros',
                      'Phalaris aquatica',
                      'Juncus bufonius',
                      'Isolepis carinata',
                      'Eleocharis macrostachya',
                      'Juncus occidentalis',
                      'Juncus phaeocephalus',
                      'Danthonia californica',
                      'Hordeum brachyantherum',
                      'Stipa pulchra',
                      'Cerastium glomeratum',
                      'Briza maxima',
                      'Vicia sativa',
                      'Dichelostemma capitatum',
                      'Holcus lanatus',
                      'Lysimachia arvensis',
                      'Briza minor',
                      'Erodium cicutarium',
                      'Hypochaeris glabra',
                      'Rumex crispus',
                      'Sonchus oleraceus',
                      'Ranunculus californicus',
                      'Rumex pulcher',
                      'Convolvulus arvensis',
                      'Lythrum hyssopifolium',
                      'Stellaria nitens',
                      'Avena barbata',
                      'Cynosurus echinatus',
                      'Poa annua',
                      'Lysimachia minima',
                      'Lupinus nanus',
                      'Brodiaea terrestris',
                      'Carex densa',
                      'Sonchus asper',
                      'Hordeum murinum',
                      'Medicago polymorpha',
                      'Silybum marianum',
                      'Carduus pycnocephalus',
                      'Brachypodium distachyon',
                      'Eschscholzia californica',
                      'Bromus carinatus',
                      'Linum bienne',
                      'Clarkia davyi',
                      'Senecio vulgaris',
                      'Sherardia arvensis',
                      'Calandrinia menziesii',
                      'Triphysaria pusilla',
                      'Gamochaeta ustulata',
                      'Cirsium vulgare',
                      'Silene gallica',
                      'Juncus patens',
                      'Carex tumulicola',
                      'Oxalis pilosa',
                      'Sisyrinchium bellum',
                      'Spergula arvensis',
                      'Epilobium brachycarpum',
                      'Hirschfeldia incana',
                      'Raphanus sativus',
                      'Avena fatua',
                      'Soliva sessilis',
                      'Epilobium ciliatum',
                      'Nassella lepida')

#  List of speices that need to be called separately:
#                       'Erigeron spp'
#                       'Carex spp',
#                       'Briza sp',
#                       'Madia sp',
#                       'Rumex spp',
#                       'Vicia spp',
#                       'Bromus sp',
#                       'Brassica species',
#                       'Brodiaea species',

all_species_data <- occ(query = all_species_list, from = "gbif", has_coords = TRUE, limit = 100000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))


```

```{r clean mcLann species data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
all_species_data <- fixnames(all_species_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_species_data_df <- occ2df(all_species_data)

# We need to clean the coordinates, as it appears some are in the ocean, and there might be some that are just incorrect. To do this, we will use a coordinate cleaner package.
flags <- clean_coordinates(x = all_species_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most tests are on by default

# Most of our flagged coordinates are coastal/in the ocean. Therefore, we'll remove them from our database.
# keep only the non-flagged data
all_species_data_df <- all_species_data_df[flags$.summary,] %>%
  rename("species.name" = "name")
```

Because there are three genera found in the plots could reflect any number of species known from the site, we fill collect gbif data for all of these.

##### Bromus
```{r elk bromus query GBIF}
all_bromus <- c('Bromus carinatus',
                'Bromus hordeaceus',
                'Bromus diandrus',
                'Avena barbata')

all_bromus_data <- occ(query = all_bromus, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean elk bromus data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_bromus_data <- fixnames(all_bromus_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_bromus_data_df <- occ2df(all_bromus_data)

flags <- clean_coordinates(x = all_bromus_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_bromus_data_df <- all_bromus_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_bromus_data_df <- all_bromus_data_df %>%
  mutate(species.name = "Bromus sp")
```

##### Brassica
```{r bromus query GBIF}
all_brassica <- c('Brassica rapa',
                  'Brassica nigra',
                  'Hirschfeldia incana')

all_brassica_data <- occ(query = all_brassica, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean bromus data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_brassica_data <- fixnames(all_brassica_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_brassica_data_df <- occ2df(all_brassica_data)

flags <- clean_coordinates(x = all_brassica_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_brassica_data_df <- all_brassica_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_brassica_data_df <- all_brassica_data_df %>%
  mutate(species.name = "Brassica species")
```

##### Rumex
```{r  rumex query GBIF}
all_rumex <- c('Rumex acetosella',
                'Rumex crispus',
                'Rumex pulcher')
         
all_rumex_data <- occ(query = all_rumex, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean  rumex data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_rumex_data <- fixnames(all_rumex_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_rumex_data_df <- occ2df(all_rumex_data)

flags <- clean_coordinates(x = all_rumex_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_rumex_data_df <- all_rumex_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_rumex_data_df <- all_rumex_data_df %>%
  mutate(species.name = "Rumex spp")
```


##### Erigeron spp

```{r Erigeron query GBIF}
all_erigeron <- c('Erigeron canadensis',
                  'Erigeron sumatrensis')
         
all_erigeron_data <- occ(query = all_erigeron, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Erigeron data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_erigeron_data <- fixnames(all_erigeron_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_erigeron_data_df <- occ2df(all_erigeron_data)

flags <- clean_coordinates(x = all_erigeron_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_erigeron_data_df <- all_erigeron_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_erigeron_data_df <- all_erigeron_data_df %>%
  mutate(species.name = "Erigeron spp")
```

##### 'Carex spp'
```{r Carex query GBIF}
all_carex <- c('Carex tumulicola',
                  'Carex densa')
         
all_carex_data <- occ(query = all_carex, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Carex data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_carex_data <- fixnames(all_carex_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_carex_data_df <- occ2df(all_carex_data)

flags <- clean_coordinates(x = all_carex_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_carex_data_df <- all_carex_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_carex_data_df <- all_carex_data_df %>%
  mutate(species.name = "Carex spp")
```

##### 'Briza sp'
```{r Briza query GBIF}
all_briza <- c('Briza maxima',
               'Briza minor')
         
all_briza_data <- occ(query = all_briza, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Briza data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_briza_data <- fixnames(all_briza_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_briza_data_df <- occ2df(all_briza_data)

flags <- clean_coordinates(x = all_briza_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_briza_data_df <- all_briza_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_briza_data_df <- all_briza_data_df %>%
  mutate(species.name = "Briza sp")
```

##### 'Madia sp'
```{r Madia query GBIF}
all_madia <- c('Madia sativa',
                  'Madia elegans')
         
all_madia_data <- occ(query = all_madia, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Madia data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_madia_data <- fixnames(all_madia_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_madia_data_df <- occ2df(all_madia_data)

flags <- clean_coordinates(x = all_madia_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_madia_data_df <- all_madia_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_madia_data_df <- all_madia_data_df %>%
  mutate(species.name = "Madia sp")
```

##### 'Vicia spp'
```{r Carex query GBIF}
all_vicia <- c('Carex tumulicola',
                  'Carex densa')
         
all_vicia_data <- occ(query = all_vicia, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Carex data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_vicia_data <- fixnames(all_vicia_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_vicia_data_df <- occ2df(all_vicia_data)

flags <- clean_coordinates(x = all_vicia_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_vicia_data_df <- all_vicia_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_vicia_data_df <- all_vicia_data_df %>%
  mutate(species.name = "Vicia spp")
```

##### 'Brodiaea species'
```{r Carex query GBIF}
all_brodiaea <- c('Brodiaea terrestris',
                  'Brodiaea elegans')
         
all_brodiaea_data <- occ(query = all_brodiaea, from = "gbif", has_coords = TRUE, limit = 10000,
                gbifopts = list(hasGeospatialIssue = FALSE, country = "US",
                                stateProvince = "California"))
```

```{r clean Carex data}
## we can also replace all synonym taxon names with one (cuts down the number of plotted dot colors)
## we should use the scrubr version of the function in the future
all_brodiaea_data <- fixnames(all_brodiaea_data, how = "query")

## The occ2df function should convert the data to a dataframe format
all_brodiaea_data_df <- occ2df(all_brodiaea_data)

flags <- clean_coordinates(x = all_brodiaea_data_df, lon = "longitude", lat = "latitude",
                          species = "name",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "seas", "zeros", "duplicates")) # most test are on by default

# remove flagged records
all_brodiaea_data_df <- all_brodiaea_data_df[flags$.summary,]

# change all datapoints to the "species sp" format that's used in the og dataframe, calc CTI
all_brodiaea_data_df <- all_brodiaea_data_df %>%
  mutate(species.name = "Brodiaea species")
```

### Combine with and without epithet dataframes and remove extra objects

```{r recombine car with epithet with non-epithet data}
all_species_data_df <- full_join(all_species_data_df, all_bromus_data_df)
all_species_data_df <- full_join(all_species_data_df, all_brassica_data_df)
all_species_data_df <- full_join(all_species_data_df, all_rumex_data_df)
all_species_data_df <- full_join(all_species_data_df, all_erigeron_data_df)
all_species_data_df <- full_join(all_species_data_df, all_carex_data_df)
all_species_data_df <- full_join(all_species_data_df, all_briza_data_df)
all_species_data_df <- full_join(all_species_data_df, all_madia_data_df)
all_species_data_df <- full_join(all_species_data_df, all_vicia_data_df)
all_species_data_df <- full_join(all_species_data_df, all_brodiaea_data_df)


# and we'll save the angelo gbif spatial data
write_rds(all_species_data_df, "Analysis/GBIF_Data/AllKaren_GBIF.rds")
```


### Combine occurance data with climate data

This method of data extraction follows the tutorial on the jcoliver webpage. 

```{r based on oikos page, cache = TRUE}
xy <- all_species_data_df[,c("longitude","latitude")]
sp_xy <- all_species_data_df[c("species.name", "longitude","latitude")]

env <- getData('worldclim', var='bio', res=2.5)

plot(env, 1, main=NULL, axes=FALSE)
points(xy, col='blue', pch=20)

xy_bio <- raster::extract(env, xy)

```

Reference list for climate variables:
 * BIO1 = Annual Mean Temperature (C * 10) -- (i.e. 231 represents 23.1 C)
 * BIO2 = Mean Diurnal Range (Mean of monthly (max temp  min temp)) 
 * BIO3 = Isothermality (BIO2/BIO7) (* 100)
 * BIO4 = Temperature Seasonality (standard deviation *100)
 * BIO5 = Max Temperature of Warmest Month (C * 10)
 * BIO6 = Min Temperature of Coldest Month (C * 10)
 * BIO7 = Temperature Annual Range (BIO5-BIO6)
 * BIO8 = Mean Temperature of Wettest Quarter (C * 10)
 * BIO9 = Mean Temperature of Driest Quarter (C * 10)
 * BIO10 = Mean Temperature of Warmest Quarter (C * 10)
 * BIO11 = Mean Temperature of Coldest Quarter (C * 10)
 * BIO12 = Annual Precipitation (mm)
 * BIO13 = Precipitation of Wettest Month (mm)
 * BIO14 = Precipitation of Driest Month (mm)
 * BIO15 = Precipitation Seasonality (Coeficent of Variation) 
 * BIO16 = Precipitation of Wettest Quarter (mm)
 * BIO17 = Precipitation of Driest Quarter (mm)
 * BIO18 = Precipitation of Warmest Quarter (mm)
 * BIO19 = Precipitation of Coldest Quarter (mm)
 
```{r mean species temp, echo=FALSE}
all_species_clim <- cbind(xy_bio, sp_xy) 

all_species_clim <- all_species_clim %>%
  group_by(species.name) %>%
  summarise(meant = mean(bio1, na.rm = TRUE),
         meanp = mean(bio12, na.rm = TRUE)) %>%
  mutate(meanT = meant/10) %>%
  dplyr::select(-meant)
```

### Combine clim data with original relative data
```{r angelo rel and clim data join}
all_karen_CTI <- left_join(rel_all, all_species_clim, by = "species.name") 

write.csv(all_karen_CTI, "Analysis/CTI Data/All_Karen_08_2020.csv")
```

# CTI Graphs

```{r}

all_karen_CTI <- all_karen_CTI %>%
  filter(species.name != "Dung",
         species.name != "Oxalis pilosa",
         freq != "CLA", freq != "CLB", freq != "CLC", freq != "NM", freq != "CL2", freq != "CL3")

CTI_plot_means <- all_karen_CTI %>%
  group_by(site, plot, freq, treat, year) %>%
  summarise(meanCTI = mean(meanT),
            seCTI = std.error(meanT),
            meanCPI = mean(meanp),
            seCPI = std.error(meanp))

CTI_site_means <- all_karen_CTI %>%
  group_by(site, freq, year) %>%
  summarise(meanCTI = mean(meanT),
            seCTI = std.error(meanT),
            meanCPI = mean(meanp),
            seCPI = std.error(meanp))

grid.arrange(
ggplot(CTI_site_means, aes(x = year, y = meanCTI, color = site)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = (meanCTI - seCTI), ymax = (meanCTI + seCTI))) +
  labs(title = "All sites - Community Temperature Index",
       x = "Year",
       y = "CTI") +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(site~freq, scales = "free_x", ncol = 3) # + geom_smooth(method = "lm")
,
ggplot(CTI_site_means, aes(x = year, y = meanCPI, color = site)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = (meanCPI - seCPI), ymax = (meanCPI + seCPI))) +
  labs(title = "All sites - Community Precipitation Index",
       x = "Year",
       y = "CPI") +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(site~freq, scales = "free_x", ncol = 3) # + geom_smooth(method = "lm")
, ncol=2
)

```

