---
title: "Supplementary figures"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/setup.R")
```

# Occurrence data sources

Compare BIEN, CCH, GBIF, iNat occurrence datasets.

```{r occ-data-tab}
source("scripts/compare-occ-data.R")
knitr::kable(occ_samp_tbl, caption = "Comparison of occurrence datasets.")
```

```{r occ-data-fig, fig.cap="Occurrence data comparison from multiple sources in study area"}
occ_comp_gg
rm(list = setdiff(ls(), "path_ls"))
```

# Climate data sources

Compare CHELSA and PRISM climate data at GBIF occurrence locations, by making scatter plots with correlation.

```{r clim-data-fig, fig.cap="Climate data comparison from multiple sources shows high consistency in study area."}
source("scripts/compare-clim-data.R")
chelsa_prism_gg / chelsa_terraclim_gg
rm(list = setdiff(ls(), "path_ls"))
```

# Climate variable correlations

TODO: different bioclim variables are highly correlated

# Climate niche summary statistics

The mean, median, upper, lower temp and precip for all species are highly correlated.

```{r tmp-niche-fig, fig.cap="Species temperature niche statistics are highly correlated"}
source("scripts/compare-niche-stat.R")
niche_tmp_gg
```

```{r ppt-niche-fig, fig.cap="Species precipitation niche statistics are highly correlated"}
niche_ppt_gg
rm(list = setdiff(ls(), "path_ls"))
```

# Spatial thinning occurrence-climate records

```{r thin-tmp-fig, fig.cap="Species temperature niche statistics (°C)"}
source("scripts/compare-sp-thin.R")
thin_tmp_gg
```

```{r thin-ppt-fig, fig.cap="Species precipitation niche statistics (mm)"}
thin_ppt_gg
rm(list = setdiff(ls(), "path_ls"))
```

# Individual species climate niches

Generate a single species per page and assembly into a multi-page plot. Left geo-space, right clim-space.

```{r}
source("scripts/plot-ind-spp.R")
```

Show two example species.

```{r cool-sp-fig, fig.cap="Example species *Danthonia californica*"}
niche_gg["Danthonia californica"]
```

```{r warm-sp-fig, fig.cap="Example species *Stipa pulchra*"}
niche_gg["Stipa pulchra"]
```

Print all species into a multi-page PDF.

```{r, eval=FALSE}
# runtime ~= 4 min
pdf("figures/species-climate-niche.pdf", width = 8, height = 8 * .618)
print(niche_gg)
dev.off()
```

The file can be viewed at this [link](https://zhulab.ucsc.edu/projects/grassland/figures/species-climate-niche.pdf)

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

# Site info

Experimental and observational site information in table and map.

```{r}
site_sf <- read_rds(path_ls$geo_site) %>%
  filter(abbr %in% c("angelo", "carrizo", "elkhorn", "jasper", "mclann", "morganterritory", "pleasantonridge", "sunol", "swanton", "ucsc", "vascocaves")) %>%
  arrange(abbr) %>%
  add_column(lab = c(LETTERS[1:4], "E/F", LETTERS[7:12]))

site_sf %>%
  as_tibble() %>% # drop geometry
  knitr::kable()
```

Get grassland coverage. Land cover type data on 0.05 degree resolution maunually downloaded from https://lpdaac.usgs.gov/products/mcd12c1v006/.

```{r}
cfp_sf <- st_read(path_ls$geo_cfp) %>%
  filter(
    NAME == "California Floristic Province",
    Type == "hotspot area"
  )

sds <- terra::sds(paste0(path_ls$geo_grass, "MCD12C1.A2020001.006.2021362215328.hdf"))
lc_ras <- raster::raster(sds[1])
raster::extent(lc_ras) <- c(-180, 180, -90, 90)
raster::projection(lc_ras) <- sp::CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
lc_ras <- raster::flip(lc_ras)
lc_ras <- raster::crop(lc_ras, raster::extent(cfp_sf))
lc_ras <- raster::mask(lc_ras, cfp_sf)

grass_ras <- lc_ras
grass_ras[grass_ras != 10] <- NA # User guide at https://lpdaac.usgs.gov/documents/101/MCD12_User_Guide_V6.pdf
grass_ras[grass_ras == 10] <- 1
grass_df <- raster::as.data.frame(grass_ras, xy = T) %>%
  drop_na()
```

```{r fig.cap="Map of experimental and observational study sites"}
set.seed(618)
ggplot() +
  geom_sf(
    data = rnaturalearth::ne_states(
      country = c("Mexico", "United States of America"),
      returnclass = "sf"
    ),
    fill = NA,
    color = alpha("black", .1)
  ) +
  geom_sf(
    data = cfp_sf,
    color = alpha("black", .3),
    fill = alpha("white", .1)
  ) +
  geom_raster(
    data = grass_df,
    aes(x = x, y = y),
    fill = "yellow green"
  ) +
  geom_sf(data = site_sf, color = "red") +
  ggrepel::geom_label_repel(
    data = site_sf,
    mapping = aes(
      label = lab, # name,
      geometry = geometry
    ),
    stat = "sf_coordinates",
    size = 3,
    color = "red",
    fill = NA,
    min.segment.length = 0,
    label.padding = unit(.1, "lines"),
    label.size = NA
  ) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
  scale_x_continuous(breaks = c(-125, -120, -115)) +
  scale_y_continuous(breaks = c(30, 35, 40))
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

# Site climate change

TODO: get PRISM/CHELSA data on six sites and plot climate change; may compare with CTI and CPI changes
