# Community shift analyses

Let us begin with the theory of moment method estimation, which is a generalization of the CTI/CPI approach. For location $i$, species $j$, denote species abundance (point intercept or percent cover) as $A_{ij}$, and species trait (e.g., optimal temperature) as $T_j$. At the community level, the expected trait (i.e., first moment), which is equivalent to CTI/CPI, is

$$
\mathrm{E}(T_i) = \bar{T_i} = \frac{\sum_j A_{ij}T_j}{\sum_j A_{ij}}.
$$

Furthermore, the trait variance (i.e., second central moment), using the variance-expectation formula, is

$$
\mathrm{Var}(T_i) = \mathrm{E}[T_i^2] - \mathrm{E}[T_i]^2 = \frac{\sum_j A_{ij}T_j^2}{\sum_j A_{ij}} - \bar{T_i}^2.
$$

A pair of $\{\mathrm{E}(T_i), \mathrm{Var}(T_i)\}$ characterizes the trait values under the distribution of species abundance. Operationally, we might use a pair of mean and standard deviation, $\{ \mu(T_i) = \mathrm{E}(T_i), \sigma(T_i) =  \sqrt{\mathrm{Var}(T_i)}\}$, so that they are comparable with the same unit. 

Setup and load packages.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, eval = TRUE)
```

```{r}
library(tidyverse)
```

Load niche estimates.

```{r}
niche_tbl <- read_rds("data/climate/niche_estimates.rds")
```

## Experimental data

Load all experimental community data. Join species climate niche estimates with community data, then aggregate by site, year, plot, treatment, calculate community-averaged temp and precip, weighted by abundance.

```{r}
exp_com_tbl <- read_rds("data/community/all-experimental-data.rds") %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot, treat) %>%
  summarize(
    tmean_com_mean = sum(abund * tmean_occ_mean) / sum(abund),
    tmean_com_var = sum(abund * tmean_occ_mean^2) / sum(abund) - tmean_com_mean^2,
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund),
    ppt_com_var = sum(abund * ppt_occ_mean^2) / sum(abund) - ppt_com_mean^2,
    vpdmax_com_mean = sum(abund * vpdmax_occ_mean) / sum(abund),
    vpdmax_com_var = sum(abund * vpdmax_occ_mean^2) / sum(abund) - vpdmax_com_mean^2
  )
```

### JRGCE 

Community-level mean CTI in warming vs. ambient plots.

```{r, fig.cap = "Compare mean(CTI) in ambient vs. elevated warming treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "T___")
  ) %>%
  ggplot(aes(year, tmean_com_mean, col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  labs(
    x = "Year", y = "Community Temperature Index (CTI, degC)",
    color = "Warming treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

Community-level CTI standard deviation in warming vs. ambient plots.

```{r, fig.cap = "Compare sd(CTI) in ambient vs. elevated warming treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "T___")
  ) %>%
  ggplot(aes(year, sqrt(tmean_com_var), col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  labs(
    x = "Year", y = "Community Temperature Index (CTI, degC)",
    color = "Warming treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

Community-level mean CPI in watered vs. ambient plots.

```{r, fig.cap = "Compare mean(CPI) in ambient vs. elevated adding water treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "_P__")
  ) %>%
  ggplot(aes(year, ppt_com_mean, col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)",
    color = "Watering treatment"
  ) +
  scale_color_manual(values = c("black", "blue"))
```

Community-level CPI standard deviation in watered vs. ambient plots.

```{r, fig.cap = "Compare sd(CPI) in ambient vs. elevated adding water treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "_P__")
  ) %>%
  ggplot(aes(year, sqrt(ppt_com_var), col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)",
    color = "Watering treatment"
  ) +
  scale_color_manual(values = c("black", "blue"))
```

For both warming and watering treatments, show replicates as raw data.

```{r, fig.cap="Plot (replicate) level community weighted temperature and precipitation indexes from ambient and elevated temp and precip treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "T___", "_P__", "TP__")
  ) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(tmean_com_mean, ppt_com_mean, group = interaction(treat, year))) +
  geom_point(aes(color = year), alpha = .5) +
  # stat_ellipse() +
  facet_wrap(~treat) +
  labs(
    x = "Community Temperature Index (CTI, degC)",
    y = "Community Precipitation Index (CPI, mm)",
    col = "Year"
  ) +
  scale_color_viridis_d()
```

Because of outliers, summarize by median (same as boxplot, not mean) values of replicates from the same treatment.

```{r, fig.cap="Median community weighted temperature and precipitation indexes from ambient and elevated temp and precip treatments over years"}
exp_com_tbl %>%
  filter(
    site == "JRGCE",
    treat %in% c("____", "T___", "_P__", "TP__")
  ) %>%
  group_by(year, treat) %>%
  summarize(tmean_com_mean = median(tmean_com_mean), ppt_com_mean = median(ppt_com_mean)) %>%
  # mutate(year = as.factor(year)) %>%
  ggplot(aes(tmean_com_mean, ppt_com_mean, col = year)) +
  geom_path() +
  facet_wrap(~treat) +
  labs(
    x = "Community Temperature Index (CTI, degC)",
    y = "Community Precipitation Index (CPI, mm)",
    col = "Year"
  ) +
  scale_color_viridis_c()
```

### Santa Cruz IDE

Plot (replicate) level mean CPI in ambient vs. drought treatments.

```{r, fig.cap="Compare mean(CPI) in ambient vs. drought treatments over years"}
exp_com_tbl %>%
  filter(site %in% c("IDE Arboretum", "IDE Marshall Field", "IDE YLR")) %>%
  ggplot(aes(year, ppt_com_mean, col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  facet_wrap(~site) +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)",
    color = "Drought treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

Michael said Marshall Field has lots of native species, while Arboretum and YLR are highly invaded (2/16/22). That might explain the difference among those sites. 

Plot (replicate) level CPI standard deviation in ambient vs. drought treatments.

```{r, fig.cap="Compare sd(CPI) in ambient vs. drought treatments over years"}
exp_com_tbl %>%
  filter(site %in% c("IDE Arboretum", "IDE Marshall Field", "IDE YLR")) %>%
  ggplot(aes(year, sqrt(ppt_com_var), col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  facet_wrap(~site) +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)",
    color = "Drought treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

Use mean VPDmax to replace total precip.

```{r, fig.cap="Compare community VPDmax in ambient vs. drought treatments over years"}
exp_com_tbl %>%
  filter(site %in% c("IDE Arboretum", "IDE Marshall Field", "IDE YLR")) %>%
  ggplot(aes(year, vpdmax_com_mean, col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  facet_wrap(~site) +
  labs(
    x = "Year", y = "Community VPDmax Index (CVI, hPa)",
    color = "Drought treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

Use VPDmax standard deviation to replace total precip.

```{r, fig.cap="Compare community VPDmax in ambient vs. drought treatments over years"}
exp_com_tbl %>%
  filter(site %in% c("IDE Arboretum", "IDE Marshall Field", "IDE YLR")) %>%
  ggplot(aes(year, sqrt(vpdmax_com_var), col = treat, group = interaction(treat, year))) +
  geom_boxplot() +
  facet_wrap(~site) +
  labs(
    x = "Year", y = "Community VPDmax Index (CVI, hPa)",
    color = "Drought treatment"
  ) +
  scale_color_manual(values = c("black", "red"))
```

## Observational data

Load all experimental community data. Join species climate niche estimates with community data, then aggregate by site, year, plot, and calculate community-averaged temp and precip, weighted by abundance.

```{r}
obs_com_tbl <- read_rds("data/community/all-observational-data.rds") %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot) %>%
  summarize(
    tmean_com_mean = sum(abund * tmean_occ_mean) / sum(abund),
    tmean_com_var = sum(abund * tmean_occ_mean^2) / sum(abund) - tmean_com_mean^2,
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund),
    ppt_com_var = sum(abund * ppt_occ_mean^2) / sum(abund) - ppt_com_mean^2,
    vpdmax_com_mean = sum(abund * vpdmax_occ_mean) / sum(abund),
    vpdmax_com_var = sum(abund * vpdmax_occ_mean^2) / sum(abund) - vpdmax_com_mean^2
  )
```

CTI community-level mean in several sites increases over years, suggesting communities are becoming "warmer," i.e., thermophilization.

```{r, fig.cap = "Compare mean(CTI) across sites over years"}
obs_com_tbl %>%
  ggplot(aes(year, tmean_com_mean, group = year)) +
  geom_boxplot() +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Community Temperature Index (CTI, degC)"
  )
```

CTI community-level standard deviation.

```{r, fig.cap = "Compare sd(CTI) across sites over years"}
obs_com_tbl %>%
  ggplot(aes(year, sqrt(tmean_com_var), group = year)) +
  geom_boxplot() +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Community Temperature Index (CTI, degC)"
  )
```

CPI community-level mean in several sites decreases over years, suggesting communities are becoming "drier," i.e., mesophilization.

```{r, fig.cap = "Compare mean(CPI) across sites over years"}
obs_com_tbl %>%
  ggplot(aes(year, ppt_com_mean, group = year)) +
  geom_boxplot() +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)"
  )
```

CPI community-level standard deviation.

```{r, fig.cap = "Compare mean(CPI) across sites over years"}
obs_com_tbl %>%
  ggplot(aes(year, sqrt(ppt_com_var), group = year)) +
  geom_boxplot() +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Community Precipitation Index (CPI, mm)"
  )
```
