# Manuscript display items

Setup and load packages.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, eval = TRUE)
```

```{r}
library(tidyverse)
library(sf)
library(patchwork)
theme_set(ggpubr::theme_pubclean())
```

## Main text

Figures and tables in the main text.

### Climate niche estimation

Read in GBIF-CHELSA, and niche estimate data.

```{r}
gbif_chelsa_sf <- read_rds("data/climate/climate-gbif-2022-05-10.rds") %>%
  select(geometry, key, species, tmp = chelsa_tmp, ppt = chelsa_ppt, vpd = chelsa_vpd) %>%
  st_as_sf(crs = "+proj=longlat +datum=WGS84 +no_defs")

niche_tbl <- read_rds("data/occurrence/niche-estimates-cfp-2022-05-10.rds")
```

Choose two example species and assign colors. From Justin's email 5/10/22:

> Are you aiming for natives specifically? I think Stipa pulchra works well for the dry, Festuca perennis is also commonly known since its a weed, in the same area. 
> 
> For wet species I don't recommend Allium, I might recommend Danthonia californica, or Juncus bufonis (Juncus are rushes and also typically though of as 'wetter' or cooler species). For a non-native if desired - Holcus lanatus would be a good choice.
>
> Danthonia californica looks like its in a good spot, but I'm sure how well known it is - Juncus bufonis could be good - most people would know Juncus typically likes 'wet and cooler' but it is not in as an extreme location.
>
> Note: Juncus bufonis should be Juncus bufonius.

Michael confirmed Danthonia californica and Stipa pulchra are two important native species in California on 5/16/22.

```{r}
cool_species <- "Danthonia californica" # "Juncus bufonius"
warm_species <- "Stipa pulchra" # "Festuca perennis"

gbif_chelsa_sf <- gbif_chelsa_sf %>%
  mutate(species_type = case_when(
    species == cool_species ~ "cool",
    species == warm_species ~ "warm",
    TRUE ~ "other"
  ))

niche_tbl <- niche_tbl %>%
  mutate(species_type = case_when(
    species == cool_species ~ "cool",
    species == warm_species ~ "warm",
    TRUE ~ "other"
  ))
```

Create three panels separately (fast without rendering).

```{r}
# set up plot params using alphabetical species type names
color_vec <- c(cool = "blue", other = gray(.9), warm = "red")
shape_vec <- c(cool = 19, other = 20, warm = 19)
alpha_vec <- c(cool = .6, other = .005, warm = .6)

occ_geog_gg <-
  ggplot() +
  geom_sf(
    data = rnaturalearth::ne_states(
      country = c("Mexico", "United States of America"),
      returnclass = "sf"
    ),
    fill = NA,
    color = gray(.25)
  ) +
  geom_sf(
    data = gbif_chelsa_sf,
    aes(color = species_type, shape = species_type, alpha = species_type)
  ) +
  scale_color_manual(values = color_vec) +
  scale_shape_manual(values = shape_vec) +
  scale_alpha_manual(values = alpha_vec) +
  coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
  scale_x_continuous(breaks = c(-125, -120, -115)) +
  scale_y_continuous(breaks = c(30, 35, 40)) +
  labs(x = "Longitude", y = "Latitude") +
  guides(color = "none", shape = "none", alpha = "none")

occ_clim_gg <-
  ggplot(data = gbif_chelsa_sf, mapping = aes(tmp, ppt)) +
  geom_point(
    aes(color = species_type, shape = species_type, alpha = species_type)
  ) +
  scale_color_manual(values = color_vec) +
  scale_shape_manual(values = shape_vec) +
  scale_alpha_manual(values = alpha_vec) +
  labs(x = "Mean annual temperature (°C)", y = "Annual precipitation (mm)") +
  guides(color = "none", shape = "none", alpha = "none")

niche_gg <-
  ggplot(
    data = niche_tbl,
    mapping = aes(
      x = tmp_occ_mean,
      y = ppt_occ_mean,
      xmin = tmp_occ_mean - tmp_occ_sd / sqrt(occ_n),
      xmax = tmp_occ_mean + tmp_occ_sd / sqrt(occ_n),
      ymin = ppt_occ_mean - ppt_occ_sd / sqrt(occ_n),
      ymax = ppt_occ_mean + ppt_occ_sd / sqrt(occ_n),
      label = species,
      color = species_type
    )
  ) +
  geom_point() +
  geom_errorbar() +
  geom_errorbarh() +
  scale_color_manual(values = c(cool = "blue", other = gray(.5), warm = "red")) +
  scale_x_continuous(expand = expansion(mult = .1)) + # expand padding to show label
  labs(x = "Mean annual temperature (°C)", y = "Annual precipitation (mm)") +
  guides(color = "none") +
  geom_label(
    data = niche_tbl %>% filter(species_type %in% c("cool", "warm")),
    fill = NA,
    fontface = "italic",
    label.padding = unit(.5, "lines"),
    label.size = 0,
    vjust = "outward", hjust = "outward"
  )
```

Combine three panels.

```{r}
fig1_gg <- occ_geog_gg + occ_clim_gg + niche_gg +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = "
  ABB
  ABB
  CCC
  CCC
  CCC
  ")
```

Render (slow) and save.

```{r, fig.width=7.5, fig.height=7.5*1.168, fig.cap="Grassland species occur along geographic and climatic gradients to estimate ecological niches. (A) Occurrence of grassland species in the California Floristic Province, retrieved from the GBIF data (*n* = 367,657 records, gray points), highlighting two example species *Danthonia californica* in the north (*n* = 894 records, blue points) and *Stipa pulchra* in the south (*n* = 481 records, red points). (B) Species occurrence in climate space of mean annual temperature (°C) and annual precipitation (mm), retrieved from the CHELSA data (gray), highlighting the same two example species *Danthonia californica* under cool and wet climates (blue) and *Stipa pulchra* under warm and dry climates (red). (C) Ecological niche estimates from average temperature and precipitation for each species (*n* = 182 species, mean ± standard error, gray), highlighting the two species (blue and red)."}
fig1_gg

# ggsave(
#   plot = fig1_gg,
#   filename = str_c(
#     "figures/fig1-",
#     word(cool_species, 1) %>% tolower(),
#     "-",
#     word(warm_species, 1) %>% tolower(),
#     ".png"
#   ),
#   width = 7.5,
#   height = 7.5 * 1.618
# )
```

### Community shift analyses

Read in experimental and observational data.

```{r}
exp_com_tbl <- read_rds("data/community/all-experimental-data.rds") %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot, treat) %>%
  summarize(
    tmp_com_mean = sum(abund * tmp_occ_mean) / sum(abund),
    tmp_com_var = sum(abund * tmp_occ_mean^2) / sum(abund) - tmp_com_mean^2,
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund),
    ppt_com_var = sum(abund * ppt_occ_mean^2) / sum(abund) - ppt_com_mean^2,
    vpd_com_mean = sum(abund * vpd_occ_mean) / sum(abund),
    vpd_com_var = sum(abund * vpd_occ_mean^2) / sum(abund) - vpd_com_mean^2
  )

obs_com_tbl <- read_rds("data/community/all-observational-data.rds") %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot) %>%
  summarize(
    tmp_com_mean = sum(abund * tmp_occ_mean) / sum(abund),
    tmp_com_var = sum(abund * tmp_occ_mean^2) / sum(abund) - tmp_com_mean^2,
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund),
    ppt_com_var = sum(abund * ppt_occ_mean^2) / sum(abund) - ppt_com_mean^2,
    vpd_com_mean = sum(abund * vpd_occ_mean) / sum(abund),
    vpd_com_var = sum(abund * vpd_occ_mean^2) / sum(abund) - vpd_com_mean^2
  )
```

JRGCE warming treatment to CTI and CPI comparisons. 

```{r}
jrgce_gg <-
  exp_com_tbl %>%
  filter(site == "jrgce") %>%
  mutate(treat_T = str_sub(treat, start = 1L, end = 1L)) %>%
  select(site, year, plot, treat_T, tmp_com_mean, ppt_com_mean) %>%
  pivot_longer(cols = tmp_com_mean:ppt_com_mean, names_to = "com_idx_name", values_to = "com_idx_value") %>%
  mutate(com_idx_name = factor(com_idx_name,
    levels = c("tmp_com_mean", "ppt_com_mean"),
    labels = c("CTI", "CPI")
  )) %>%
  ggplot(aes(year, com_idx_value, col = treat_T, group = interaction(treat_T, year))) +
  geom_boxplot() +
  ggpubr::stat_compare_means(
    aes(group = treat_T),
    method = "wilcox.test",
    label = "p.signif", hide.ns = FALSE
  ) +
  facet_wrap(~com_idx_name,
    ncol = 1, scales = "free_y",
    strip.position = "left",
    labeller = as_labeller(c(
      CTI = "Community Temperature Index\n(CTI, °C)",
      CPI = "Community Precipitation Index\n(CPI, mm)"
    ))
  ) +
  labs(
    x = NULL, # "Year",
    y = NULL,
    color = "Warming treatment",
    title = "Jasper Ridge Global Change Experiment",
    subtitle = "<span style='color:black'>Ambient vs. </span><span style='color:red'>warming treatment</span>"
  ) +
  scale_color_manual(values = c("black", "red")) +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    legend.position = "none",
    plot.subtitle = ggtext::element_markdown()
  )
```

Observational data CTI and CPI comparisons.

```{r}
site_labeller <- as_labeller(c(
  angelo = "Angelo Coast Range Reserve",
  carrizo = "Carrizo Plain",
  elkhorn = "Elkhorn Slough",
  jasper = "Jasper Ridge Biological Preserve",
  mclann = "McLaughlin Reserve",
  # mclserp = "McLaughlin (Serpentine)",
  # swanton = "Swanton Ranch",
  ucsc = "UC Santa Cruz"
))

obs_com_tbl <-
  obs_com_tbl %>%
  filter(site != "mclserp", site != "swanton")

obs_cti_gg <-
  obs_com_tbl %>%
  ggplot(aes(year, tmp_com_mean, group = year)) +
  geom_boxplot() +
  geom_smooth(
    method = "lm", formula = y ~ x, se = FALSE,
    color = "red", lty = "dashed",
    aes(group = 1)
  ) +
  ggpubr::stat_cor(aes(group = 1, label = ..p.label..),
    p.accuracy = 0.05,
    color = "red"
  ) +
  facet_wrap(~site,
    nrow = 2,
    scales = "free_y",
    labeller = site_labeller
  ) +
  labs(
    x = NULL, # "Year",
    y = "Community Temperature Index\n(CTI, °C)"
  ) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0)
  )

obs_cpi_gg <-
  obs_com_tbl %>%
  ggplot(aes(year, ppt_com_mean, group = year)) +
  geom_boxplot() +
  geom_smooth(
    method = "lm", formula = y ~ x, se = FALSE,
    color = "red", lty = "dashed",
    aes(group = 1)
  ) +
  ggpubr::stat_cor(aes(group = 1, label = ..p.label..),
    p.accuracy = 0.05,
    color = "red"
  ) +
  facet_wrap(~site,
    nrow = 2,
    scales = "free_y",
    labeller = site_labeller
  ) +
  labs(
    x = "Year", y = "Community Precipitation Index\n(CPI, mm)"
  ) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0)
  )
```

Combine panels.

```{r}
fig2_gg <- jrgce_gg + obs_cti_gg + obs_cpi_gg +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = "
  A
  B
  C
  ")
```

Render and save.

```{r, fig.width=10, fig.height=10*1.618, fig.cap="Grassland community shifts from long-term experiments and observations. (A) Warming treatment causes communities to shift dominance to warmer (CTI, °C) and drier (CPI, mm) species in the Jasper Ridge Global Change Experiment (*n* = 200 plots, with 136 ambient plots in black, 64 warming plots in red, 1998 pre-treatment year, non-parametric two-sample Wilcoxon test, ns: *p* > 0.05, $*$: *p* ≤ 0.05, $**$: *p* ≤ 0.01). (B and C) Consistent with climate warming and drying, grassland communities shift dominance to warmer (CTI) and drier (CPI) species in six long-term observational sites across the California Floristic Province (dashed trend lines by linear regression, all significant slopes by *t*-test, *p* < 0.05)."}
fig2_gg

# ggsave(
#   plot = fig2_gg,
#   filename = "figures/fig2.png",
#   width = 10,
#   height = 10 * 1.618
# )
```

Alternative panel arrangement by observational sites.

```{r}
obs_gg_fun <- function(tbl, site_lab, cti_lab = "", cpi_lab = "", yr_lab = NULL) {
  ggplot(data = tbl, aes(year, com_idx_value, group = year)) +
    geom_boxplot() +
    geom_smooth(
      method = "lm", formula = y ~ x, se = FALSE,
      color = "red", lty = "dashed",
      aes(group = 1)
    ) +
    ggpubr::stat_cor(aes(group = 1, label = ..p.label..),
      p.accuracy = 0.05,
      color = "red"
    ) +
    facet_wrap(~com_idx_name,
      ncol = 1, scales = "free_y",
      strip.position = "left",
      labeller = as_labeller(c(
        CTI = cti_lab, # "Community Temperature Index\n(CTI, °C)",
        CPI = cpi_lab # "Community Precipitation Index\n(CPI, mm)"
      ))
    ) +
    xlim(1983, 2019) +
    labs(
      x = yr_lab, y = NULL,
      title = site_lab
    ) +
    theme(
      strip.background = element_blank(),
      strip.placement = "outside",
      plot.title = element_text(size = 11)
    )
}

obs_com_idx_tbl <- obs_com_tbl %>%
  select(site, year, plot, tmp_com_mean, ppt_com_mean) %>%
  pivot_longer(cols = tmp_com_mean:ppt_com_mean, names_to = "com_idx_name", values_to = "com_idx_value") %>%
  mutate(com_idx_name = factor(com_idx_name,
    levels = c("tmp_com_mean", "ppt_com_mean"),
    labels = c("CTI", "CPI")
  ))

angelo_gg <- obs_com_idx_tbl %>%
  filter(site == "angelo") %>%
  obs_gg_fun(site_labeller("angelo")[[1]],
    cti_lab = "CTI (°C)", cpi_lab = "CPI (mm)"
  )

carrizo_gg <- obs_com_idx_tbl %>%
  filter(site == "carrizo") %>%
  obs_gg_fun(site_labeller("carrizo")[[1]])

elkhorn_gg <- obs_com_idx_tbl %>%
  filter(site == "elkhorn") %>%
  obs_gg_fun(site_labeller("elkhorn")[[1]])

jasper_gg <- obs_com_idx_tbl %>%
  filter(site == "jasper") %>%
  obs_gg_fun(site_labeller("jasper")[[1]],
    cti_lab = "CTI (°C)", cpi_lab = "CPI (mm)"
  )

mclann_gg <- obs_com_idx_tbl %>%
  filter(site == "mclann") %>%
  obs_gg_fun(site_labeller("mclann")[[1]],
    yr_lab = "Year"
  )

ucsc_gg <- obs_com_idx_tbl %>%
  filter(site == "ucsc") %>%
  obs_gg_fun(site_labeller("ucsc")[[1]])
```

Combine panels.

```{r}
fig3_gg <- jrgce_gg +
  angelo_gg + carrizo_gg + elkhorn_gg +
  jasper_gg + mclann_gg + ucsc_gg +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = "
  AAA
  BCD
  EFG
  ")
```

Render and save.

```{r, fig.width=10, fig.height=10*1.618, fig.cap=""}
fig3_gg

ggsave(
  plot = fig3_gg,
  filename = "figures/fig3.png",
  width = 10,
  height = 10 * 1.618
)
```

## Supplements

Figures and tables in the supplements.
