# Manuscript display items

Setup and load packages. Data path tags are

- `com`: grassland community data
- `geo`: geography data, including GIS and climate
- `occ`: occurrence data
- `sum`: summary data, e.g., niche estimates

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, eval = TRUE, warning = FALSE)

path_ls <- list(
  com_exp = "data/community/all-experimental-data.rds",
  com_obs = "data/community/all-observational-data.rds",
  geo_cfp = "data/occurrence/hotspots/hotspots_2016_1.shp",
  geo_clim = "data/climate/climate-gbif-2022-07-12.rds",
  geo_grass = "data/occurrence/grassland/",
  geo_site = "data/community/site-info.rds",
  occ_bien = "data/occurrence/bien/bien-cfp-2022-04-27.rds",
  occ_cch = "data/occurrence/cch/cch-cfp-2022-07-12.rds",
  occ_gbif = "data/occurrence/gbif/gbif-consolidated-2022-07-12.rds",
  occ_inat = "data/occurrence/inat/inat-cfp-2022-07-12.rds",
  sum_niche = "data/occurrence/niche-estimates-cfp-2022-07-12.rds",
  sum_thin = "data/occurrence/niche-estimates-cfp-2022-07-12-thin.rds"
)
```

```{r}
library(tidyverse)
library(sf)
library(patchwork)
theme_set(ggthemes::theme_few())
```

## Main text

Figures and tables in the main text.

### Climate niche estimates

Read in GBIF-CHELSA, and niche estimate data.

```{r}
gbif_chelsa_sf <- read_rds(path_ls$geo_clim) %>%
  dplyr::select(geometry, key, species, tmp = chelsa_tmp, ppt = chelsa_ppt, vpd = chelsa_vpd) %>%
  st_as_sf(crs = "+proj=longlat +datum=WGS84 +no_defs")

niche_tbl <- read_rds(path_ls$sum_niche)
```

Choose two example species and assign colors.

```{r}
cool_species <- "Danthonia californica" # Michael: important native species in California
warm_species <- "Stipa pulchra" # Susan: Stipa pulchra is the state grass, and is the subject of a lot of ecological research and restoration effort

gbif_chelsa_sf <- gbif_chelsa_sf %>%
  mutate(species_type = case_when(
    species == cool_species ~ "cool",
    species == warm_species ~ "warm",
    TRUE ~ "other"
  ))

niche_tbl <- niche_tbl %>%
  mutate(species_type = case_when(
    species == cool_species ~ "cool",
    species == warm_species ~ "warm",
    TRUE ~ "other"
  ))
```

Create three panels separately (fast without rendering).

```{r}
# set up plot params using alphabetical species type names
color_vec <- c(cool = "blue", other = gray(.9), warm = "red")
shape_vec <- c(cool = 19, other = 20, warm = 19)
alpha_vec <- c(cool = .6, other = .005, warm = .6)

occ_geog_gg <-
  ggplot() +
  geom_sf(
    data = rnaturalearth::ne_states(
      country = c("Mexico", "United States of America"),
      returnclass = "sf"
    ),
    fill = NA,
    color = gray(.25)
  ) +
  geom_sf(
    data = gbif_chelsa_sf,
    aes(color = species_type, shape = species_type, alpha = species_type)
  ) +
  scale_color_manual(values = color_vec) +
  scale_shape_manual(values = shape_vec) +
  scale_alpha_manual(values = alpha_vec) +
  coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
  scale_x_continuous(breaks = c(-125, -120, -115)) +
  scale_y_continuous(breaks = c(30, 35, 40)) +
  labs(x = "Longitude", y = "Latitude") +
  guides(color = "none", shape = "none", alpha = "none")

occ_clim_gg <-
  ggplot(data = gbif_chelsa_sf, mapping = aes(tmp, ppt)) +
  geom_point(
    aes(color = species_type, shape = species_type, alpha = species_type)
  ) +
  scale_color_manual(values = color_vec) +
  scale_shape_manual(values = shape_vec) +
  scale_alpha_manual(values = alpha_vec) +
  labs(x = "Mean annual temperature (°C)", y = "Mean annual precipitation (mm)") +
  guides(color = "none", shape = "none", alpha = "none")

clim_niche_gg <-
  ggplot(
    data = niche_tbl %>%
      filter(
        occ_n > 1,
        !str_detect(species, "DUMMY")
      ),
    mapping = aes(
      x = tmp_occ_mean,
      y = ppt_occ_mean,
      xmin = tmp_occ_mean - tmp_occ_sd / sqrt(occ_n),
      xmax = tmp_occ_mean + tmp_occ_sd / sqrt(occ_n),
      ymin = ppt_occ_mean - ppt_occ_sd / sqrt(occ_n),
      ymax = ppt_occ_mean + ppt_occ_sd / sqrt(occ_n),
      label = species,
      color = species_type
    )
  ) +
  geom_point() +
  geom_errorbar() +
  geom_errorbarh() +
  scale_color_manual(values = c(cool = "blue", other = gray(.5), warm = "red")) +
  scale_x_continuous(expand = expansion(mult = .1)) + # expand padding to show label
  scale_y_continuous(expand = expansion(mult = .1)) +
  labs(x = "Mean annual temperature (°C)", y = "Mean annual precipitation (mm)") +
  guides(color = "none") +
  geom_label(
    data = niche_tbl %>% filter(species_type %in% c("cool", "warm")),
    fill = NA,
    fontface = "italic",
    label.padding = unit(.5, "lines"),
    label.size = 0,
    vjust = "outward", hjust = "outward"
  )
```

Combine three panels.

```{r}
niche_gg <- occ_geog_gg + occ_clim_gg + clim_niche_gg +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = "
  ABB
  ABB
  CCC
  CCC
  CCC
  ")
```

Render (slow) and save.

```{r, fig.width=7.5, fig.height=7.5*1.168, fig.cap="Grassland species occur along geographic and climatic gradients to estimate ecological niches. (A) Occurrence of grassland species in the California Floristic Province, retrieved from the GBIF data (*n* = 824,478 records, gray points), highlighting two example species *Danthonia californica* in the north (*n* = 916 records, blue points) and *Stipa pulchra* in the south (*n* = 3,087 records, red points). (B) Species occurrence in climate space of mean annual temperature (°C) and annual precipitation (mm), retrieved from the CHELSA data (gray), highlighting the same two example species *Danthonia californica* under cool and wet climates (blue) and *Stipa pulchra* under warm and dry climates (red). (C) Ecological niche estimates from average temperature and precipitation for each species (*n* = 372 species, mean ± standard error, gray), highlighting the two species (blue and red)."}
niche_gg
```

```{r, eval=FALSE}
ggsave(
  plot = niche_gg,
  filename = "figures/fig1-niche.png",
  width = 7.5,
  height = 7.5 * 1.618
)
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Experimental community shifts

JRGCE warming experment to CTI and CPI comparisons. 

Read in niche and experimental data.

```{r}
niche_tbl <- read_rds(path_ls$sum_niche)

exp_tbl <- read_rds(path_ls$com_exp) %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot, treat) %>%
  summarize(
    tmp_com_mean = sum(abund * tmp_occ_mean) / sum(abund),
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund)
  )
```

Side-by-side boxplots to compare CTI and CPI from ambient vs. warming treatments.

```{r}
exp_gg <-
  exp_tbl %>%
  filter(site == "jrgce") %>%
  mutate(treat_T = str_sub(treat, start = 1L, end = 1L)) %>%
  dplyr::select(site, year, plot, treat_T, tmp_com_mean, ppt_com_mean) %>%
  pivot_longer(cols = tmp_com_mean:ppt_com_mean, names_to = "com_idx_name", values_to = "com_idx_value") %>%
  mutate(com_idx_name = factor(com_idx_name,
    levels = c("tmp_com_mean", "ppt_com_mean"),
    labels = c("CTI", "CPI")
  )) %>%
  ggplot(aes(year, com_idx_value, col = treat_T, group = interaction(treat_T, year))) +
  geom_boxplot() +
  ggpubr::stat_compare_means(
    aes(group = treat_T),
    method = "wilcox.test",
    label = "p.signif", hide.ns = FALSE
  ) +
  facet_wrap(~com_idx_name,
    ncol = 1, scales = "free_y",
    strip.position = "left",
    labeller = labeller(com_idx_name = c(
      CTI = "Community Temperature Index\n(CTI, °C)",
      CPI = "Community Precipitation Index\n(CPI, mm)"
    ))
  ) +
  scale_y_continuous(expand = expansion(mult = .1)) + # expand padding to show significance tests
  labs(
    x = NULL, # "Year",
    y = NULL,
    color = "Warming treatment",
    title = "Jasper Ridge Global Change Experiment",
    subtitle = "<span style='color:black'>Ambient vs. </span><span style='color:red'>warming treatment</span>"
  ) +
  scale_color_manual(values = c("black", "red")) +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    legend.position = "none",
    plot.subtitle = ggtext::element_markdown()
  )
```

```{r, fig.width=10, fig.height=6.18, fig.cap="Grassland community shifts from long-term experiments. Warming treatment causes communities to shift dominance to warmer (CTI, °C) and drier (CPI, mm) species in the Jasper Ridge Global Change Experiment (*n* = 200 plots, with 136 ambient plots in black, 64 warming plots in red, 1998 pre-treatment year, non-parametric two-sample Wilcoxon test, ns: *p* > 0.05, $*$: *p* ≤ 0.05, $**$: *p* ≤ 0.01, $***$: *p* ≤ 0.001). "}
exp_gg
```

```{r, eval=FALSE}
ggsave(
  plot = exp_gg,
  filename = "figures/fig2-exp.png",
  width = 10,
  height = 6.18
)
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Observational community shifts

Observational data CTI and CPI comparisons, arranged by observational sites. 

Read in niche and observational data.

```{r}
niche_tbl <- read_rds(path_ls$sum_niche)

obs_tbl <- read_rds(path_ls$com_obs) %>%
  inner_join(niche_tbl, by = "species") %>%
  group_by(site, year, plot) %>%
  summarize(
    tmp_com_mean = sum(abund * tmp_occ_mean) / sum(abund),
    ppt_com_mean = sum(abund * ppt_occ_mean) / sum(abund)
  )
```

Set up site labels.

```{r}
site_vec <- c(
  angelo = "Angelo Coast",
  carrizo = "Carrizo Plain",
  elkhorn = "Elkhorn Slough",
  jasper = "Jasper Ridge",
  mclann = "McLaughlin Annual",
  mclserp = "McLaughlin Serpentine",
  pleasantonridge = "Pleasanton Ridge",
  sunol = "Sunol",
  swanton = "Swanton Ranch",
  ucsc = "UC Santa Cruz",
  vascocaves = "Vasco Caves"
)
```

Reshape data.

```{r}
obs_idx_tbl <- obs_tbl %>%
  dplyr::select(site, year, plot, tmp_com_mean, ppt_com_mean) %>%
  pivot_longer(cols = tmp_com_mean:ppt_com_mean, names_to = "com_idx_name", values_to = "com_idx_value") %>%
  mutate(com_idx_name = factor(com_idx_name,
    levels = c("tmp_com_mean", "ppt_com_mean"),
    labels = c("CTI", "CPI")
  ))

samp_size_tbl <- obs_tbl %>%
  group_by(site) %>%
  summarize(
    n_plot = n_distinct(plot),
    n_year = n_distinct(year)
  )
```

Define the plotting function.

```{r}
obs_gg_fun <- function(tbl, site_name, cti_lab = "", cpi_lab = "", yr_lab = NULL) {
  # prepare site data
  site_lab <- site_vec[site_name]
  site_tbl <- tbl %>%
    filter(site == site_name) %>%
    group_by(com_idx_name) %>%
    nest() %>%
    mutate( # lm and p value
      p_val = map(data, ~ lm(com_idx_value ~ year, data = .)) %>%
        map_dbl(~ broom::glance(.) %>% pull(p.value))
    ) %>%
    unnest(cols = data)

  # plot
  ggplot(site_tbl, aes(year, com_idx_value)) +
    geom_boxplot(aes(group = year)) +
    geom_smooth( # add lm trend line when significant
      data = . %>% filter(p_val < 0.05),
      method = "lm", se = FALSE,
      color = "red"
    ) +
    facet_wrap(~com_idx_name,
      ncol = 1, scales = "free_y",
      strip.position = "left",
      labeller = labeller(com_idx_name = c(
        CTI = cti_lab, # "Community Temperature Index\n(CTI, °C)",
        CPI = cpi_lab # "Community Precipitation Index\n(CPI, mm)"
      ))
    ) +
    expand_limits(x = range(tbl$year)) +
    labs(
      x = yr_lab, y = NULL,
      title = site_lab
    ) +
    theme(
      strip.background = element_blank(),
      strip.placement = "outside",
      plot.title = element_text(size = 11)
    )
}
```

Apply the plotting function, make individual panels, and combine them.

```{r}
obs_gg <-
  obs_gg_fun(obs_idx_tbl, "angelo", cti_lab = "CTI (°C)", cpi_lab = "CPI (mm)") +
  obs_gg_fun(obs_idx_tbl, "carrizo") +
  obs_gg_fun(obs_idx_tbl, "elkhorn") +
  obs_gg_fun(obs_idx_tbl, "jasper") +
  obs_gg_fun(obs_idx_tbl, "mclann", cti_lab = "CTI (°C)", cpi_lab = "CPI (mm)") +
  obs_gg_fun(obs_idx_tbl, "mclserp") +
  obs_gg_fun(obs_idx_tbl, "pleasantonridge") +
  obs_gg_fun(obs_idx_tbl, "sunol") +
  obs_gg_fun(obs_idx_tbl, "swanton", cti_lab = "CTI (°C)", cpi_lab = "CPI (mm)") +
  obs_gg_fun(obs_idx_tbl, "ucsc") +
  obs_gg_fun(obs_idx_tbl, "vascocaves") +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = "
  ABCD
  EFGH
  IJKL
  ")
```

Render and save.

```{r, fig.width=10, fig.height=10*1.618, fig.cap="Grassland community shifts from long-term observations. (A to K) Consistent with climate warming and drying, grassland communities shift dominance to warmer (CTI) and drier (CPI) species in 11 long-term observational sites across the California Floristic Province (red trend lines by linear regression, significant slopes by *t*-test, *p* < 0.05)."}
obs_gg
```

```{r, eval=FALSE}
ggsave(
  plot = obs_gg,
  filename = "figures/fig3-obs.png",
  width = 10,
  height = 10 * 1.618
)
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

## Supplements

Figures and tables in the supplements.

### Occurrence data sources

Compare BIEN, CCH, GBIF, iNat occurrence datasets.

```{r}
bien_tbl <- read_rds(path_ls$occ_bien) %>%
  mutate(
    dataset = "bien",
    species = queryName,
    key = as.character(key)
  ) %>%
  select(dataset, key, species, longitude, latitude)

cch_tbl <- read_rds(path_ls$occ_cch) %>%
  mutate(
    dataset = "cch",
    species = species_name,
    key = as.character(key)
  ) %>%
  select(dataset, key, species, longitude, latitude)

gbif_tbl <- read_rds(path_ls$occ_gbif) %>%
  mutate(
    dataset = "gbif",
    species = consolidatedName,
    key = as.character(key)
  ) %>%
  filter(coordinatePrecision < 0.01 | is.na(coordinatePrecision)) %>%
  filter(coordinateUncertaintyInMeters < 10000 | is.na(coordinateUncertaintyInMeters)) %>%
  filter(!coordinateUncertaintyInMeters %in% c(301, 3036, 999, 9999)) %>%
  select(dataset, key, species, longitude, latitude)

inat_tbl <- read_rds(path_ls$occ_inat) %>%
  mutate(
    dataset = "inat",
    species = consolidatedName,
    key = as.character(key)
  ) %>%
  filter(coordinatePrecision < 0.01 | is.na(coordinatePrecision)) %>%
  filter(coordinateUncertaintyInMeters < 10000 | is.na(coordinateUncertaintyInMeters)) %>%
  filter(!coordinateUncertaintyInMeters %in% c(301, 3036, 999, 9999)) %>%
  select(dataset, key, species, longitude, latitude)
```

Combine them and convert to sf.

```{r}
occ_sf <- bien_tbl %>%
  bind_rows(cch_tbl) %>%
  bind_rows(gbif_tbl) %>%
  bind_rows(inat_tbl) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) # WGS84
```

Dataset names.

```{r}
dataset_vec <- c(
  bien = "BIEN",
  cch = "eJepson (CCH)",
  gbif = "GBIF",
  inat = "iNaturalist"
)
```

Compare sample sizes across data sets.

```{r}
occ_sf %>%
  as_tibble() %>%
  group_by(dataset) %>%
  summarise(
    record_number = n() %>% format(trim = TRUE, scientific = FALSE, big.mark = ","),
    species_number = unique(species) %>% length() %>% format(trim = TRUE, scientific = FALSE, big.mark = ",")
  ) %>%
  knitr::kable(caption = "Comparison of occurrence datasets.")
```

```{r fig.cap="Occurrence data comparison from multiple sources in study area"}
set.seed(618)
prop_samp <- .1 # prop to sample and plot
ggplot() +
  geom_sf(
    data = rnaturalearth::ne_states(
      country = c("Mexico", "United States of America"),
      returnclass = "sf"
    ),
    fill = NA,
    color = alpha("black", .1)
  ) +
  # geom_sf(data = cfp_sf, fill = "white", alpha = .5) +
  geom_sf(
    data = occ_sf %>% slice_sample(prop = prop_samp), # randomly sample certain proportion, otherwise too slow to render
    color = "black", alpha = .1, size = .1
  ) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
  scale_x_continuous(breaks = c(-125, -120, -115)) +
  scale_y_continuous(breaks = c(30, 35, 40)) +
  facet_wrap(. ~ dataset,
    nrow = 1,
    labeller = labeller(dataset = dataset_vec)
  ) +
  theme(
    strip.background = element_blank(),
    # strip.text.x = element_text(hjust = 0)
  )
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Climate data sources

Compare CHELSA and PRISM climate data at GBIF occurrence locations, by making scatter plots with correlation.

```{r}
clim_gbif_tbl <- read_rds(path_ls$geo_clim) %>%
  as_tibble()

chelsa_gbif_tbl <- clim_gbif_tbl %>%
  dplyr::select(
    geometry, key, species,
    chelsa_tmp, chelsa_ppt
  ) %>%
  pivot_longer(chelsa_tmp:chelsa_ppt,
    names_to = "variable",
    values_to = "CHELSA"
  ) %>%
  mutate(variable = str_sub(variable, start = -3L, end = -1L))

prism_gbif_tbl <- clim_gbif_tbl %>%
  dplyr::select(
    geometry, key, species,
    prism_tmp, prism_ppt
  ) %>%
  pivot_longer(prism_tmp:prism_ppt,
    names_to = "variable",
    values_to = "PRISM"
  ) %>%
  mutate(variable = str_sub(variable, start = -3L, end = -1L))

terraclim_gbif_tbl <- clim_gbif_tbl %>%
  select(
    geometry, key, species,
    terraclim_tmp, terraclim_ppt
  ) %>%
  pivot_longer(terraclim_tmp:terraclim_ppt,
    names_to = "variable",
    values_to = "TerraClim"
  ) %>%
  mutate(variable = str_sub(variable, start = -3L, end = -1L))

clim_comp_tbl <- chelsa_gbif_tbl %>%
  inner_join(
    prism_gbif_tbl,
    by = c("geometry", "key", "species", "variable")
  ) %>%
  inner_join(
    terraclim_gbif_tbl,
    by = c("geometry", "key", "species", "variable")
  )
```

```{r fig.cap="Climate data comparison from multiple sources shows high consistency in study area."}
chelsa_prism_gg <- ggplot(clim_comp_tbl, aes(CHELSA, PRISM)) +
  geom_hex(bins = 100) +
  viridis::scale_fill_viridis() +
  geom_smooth(method = "lm", formula = y ~ x - 1, se = FALSE, color = "red", lty = "dashed") +
  ggpubr::stat_cor(
    aes(label = paste(..r.label.., ..p.label.., sep = "*`,`~")),
    p.accuracy = 0.05,
    color = "red"
  ) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  facet_wrap(. ~ factor(
    variable,
    levels = c("tmp", "ppt"),
    labels = c(
      "Mean annual temperature (°C)",
      "Mean annual precipitation (mm)"
    )
  ),
  scales = "free"
  ) +
  labs(
    x = "",
    y = "PRISM",
    # fill = "Occurrence record"
  ) +
  guides(fill = "none") +
  theme(
    aspect.ratio = 1, # can't do + coord_fixed() with facet_wrap(scales = "free")
    strip.background = element_blank(),
    strip.text.x = element_text(size = 12)
  )

chelsa_terraclim_gg <- ggplot(clim_comp_tbl, aes(CHELSA, TerraClim)) +
  geom_hex(bins = 100) +
  viridis::scale_fill_viridis() +
  geom_smooth(method = "lm", formula = y ~ x - 1, se = FALSE, color = "red", lty = "dashed") +
  ggpubr::stat_cor(
    aes(label = paste(..r.label.., ..p.label.., sep = "*`,`~")),
    p.accuracy = 0.05,
    color = "red"
  ) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  facet_wrap(. ~ factor(
    variable,
    levels = c("tmp", "ppt"),
    labels = c(
      "Mean annual temperature (°C)",
      "Mean annual precipitation (mm)"
    )
  ),
  scales = "free"
  ) +
  labs(
    x = "CHELSA",
    y = "TerraClim",
    # fill = "Occurrence record"
  ) +
  guides(fill = "none") +
  theme(
    aspect.ratio = 1, # can't do + coord_fixed() with facet_wrap(scales = "free")
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )

chelsa_prism_gg / chelsa_terraclim_gg
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Climate variable correlations

TODO: different bioclim variables are highly correlated

### Climate niche summary statistics

Read species climate niche estimates.

```{r}
niche_tbl <- read_rds(path_ls$sum_niche) %>%
  filter(
    occ_n > 1,
    !str_detect(species, "DUMMY")
  )
```

The mean, median, upper, lower temp and precip for all species are highly correlated.

```{r, fig.cap="Species temperature niche statistics are highly correlated"}
niche_tbl %>%
  dplyr::select(tmp_occ_mean, tmp_occ_median, tmp_occ_q05, tmp_occ_q95) %>%
  GGally::ggpairs(
    lower = list(continuous = GGally::wrap("points", alpha = 0.2)),
    title = "Species temperature niche (°C)",
    columnLabels = c("Mean", "Median", "Lower limit (5%)", "Upper limit (95%)")
  ) +
  theme(strip.background = element_blank())
```

```{r, fig.cap="Species precipitation niche statistics are highly correlated"}
niche_tbl %>%
  dplyr::select(ppt_occ_mean, ppt_occ_median, ppt_occ_q05, ppt_occ_q95) %>%
  GGally::ggpairs(
    lower = list(continuous = GGally::wrap("points", alpha = 0.2)),
    title = "Species precipitation niche (mm)",
    columnLabels = c("Mean", "Median", "Lower limit (5%)", "Upper limit (95%)")
  ) +
  theme(strip.background = element_blank())
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Spatial thinning occurrence-climate records

```{r}
thin_tbl <- read_rds(path_ls$sum_thin)
```

```{r fig.cap="Species temperature niche statistics (°C)"}
thin_tbl %>%
  select(
    tmp_mean_full, tmp_mean_thin,
    tmp_median_full, tmp_median_thin
  ) %>%
  GGally::ggpairs(
    lower = list(continuous = GGally::wrap("points", alpha = 0.2)),
    title = "Species temperature niche (°C)",
    columnLabels = c("Full mean", "Thinned mean", "Full median", "Thinned median")
  )
```

```{r fig.cap="Species precipitation niche statistics (mm)"}
thin_tbl %>%
  select(
    ppt_mean_full, ppt_mean_thin,
    ppt_median_full, ppt_median_thin
  ) %>%
  GGally::ggpairs(
    lower = list(continuous = GGally::wrap("points", alpha = 0.2)),
    title = "Species precipitation niche (mm)",
    columnLabels = c("Full mean", "Thinned mean", "Full median", "Thinned median")
  )
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Individual species climate niches

Read CFP data.

```{r}
cfp_sf <- st_read(path_ls$geo_cfp) %>%
  filter(
    NAME == "California Floristic Province",
    Type == "hotspot area"
  )
```

Read GBIF-CHELSA data.

```{r}
gbif_chelsa_sf <- read_rds(path_ls$geo_clim) %>%
  dplyr::select(geometry, key, species, tmp = chelsa_tmp, ppt = chelsa_ppt, vpd = chelsa_vpd) %>%
  st_as_sf(crs = "+proj=longlat +datum=WGS84 +no_defs")

# check species lists
sp_gbif_vec <- gbif_chelsa_sf %>%
  pull(species) %>%
  unique() %>%
  sort()

# gbif summary
tmp_rng <- range(gbif_chelsa_sf$tmp)
ppt_rng <- range(gbif_chelsa_sf$ppt)
n_occ_tot <- nrow(gbif_chelsa_sf)
```

Multi-page PDF, each for a single species (to avoid overplotting), left geo-space, right clim-space.

```{r eval=TRUE}
# multi-species plot
niche_gg <- vector(mode = "list")
for (i in seq_along(sp_gbif_vec)) {
  sp <- sp_gbif_vec[i]
  occ_sp_sf <- filter(gbif_chelsa_sf, species == sp)
  occ_sp_stat <- occ_sp_sf %>%
    as_tibble() %>%
    summarise(
      occ_n = n(),
      tmp_occ_mean = mean(tmp, na.rm = TRUE),
      tmp_occ_sd = sd(tmp, na.rm = TRUE),
      ppt_occ_mean = mean(ppt, na.rm = TRUE),
      ppt_occ_sd = sd(ppt, na.rm = TRUE)
    )

  occ_geog <-
    ggplot() +
    geom_sf(
      data = rnaturalearth::ne_states(
        country = c("Mexico", "United States of America"),
        returnclass = "sf"
      ),
      fill = NA,
      color = alpha("black", .1)
    ) +
    geom_sf(
      data = cfp_sf,
      color = alpha("black", .3),
      fill = alpha("white", .1)
    ) +
    geom_sf(data = occ_sp_sf, alpha = .1, size = .5) +
    labs(x = "Longitude", y = "Latitude", title = sp) +
    coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
    scale_x_continuous(breaks = c(-125, -120, -115)) +
    scale_y_continuous(breaks = c(30, 35, 40)) +
    theme(plot.title = element_text(face = "italic"))

  occ_clim <-
    ggplot(occ_sp_sf, aes(tmp, ppt)) +
    geom_point(alpha = .1, size = .5) +
    geom_rug(alpha = .1) +
    stat_ellipse(col = "red") +
    geom_point(
      data = occ_sp_stat,
      aes(x = tmp_occ_mean, y = ppt_occ_mean),
      shape = 3, col = "red", size = 10
    ) +
    lims(x = tmp_rng, y = ppt_rng) +
    labs(
      x = "Mean annual temperature (°C)", y = "Mean annual precipitation (mm)",
      title = bquote(italic("n") ~ "=" ~ .(format(occ_sp_stat$occ_n, big.mark = ",", trim = TRUE)))
    )

  # print(sp)
  niche_gg[[i]] <- occ_geog + occ_clim # no need to print(); will slow down
}
names(niche_gg) <- sp_gbif_vec
```

Show two example species.

```{r fig.cap="Example species *Danthonia californica*"}
niche_gg["Danthonia californica"]
```

```{r fig.cap="Example species *Stipa pulchra*"}
niche_gg["Stipa pulchra"]
```

Print all species into a multi-page PDF.

```{r, eval=FALSE}
# runtime ~= 4 min
pdf("figures/species-climate-niche.pdf", width = 8, height = 8 * .618)
print(niche_gg)
dev.off()
```

The file can be viewed at this [link](https://zhulab.ucsc.edu/projects/grassland/figures/species-climate-niche.pdf)

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Site info

Experimental and observational site information in table and map.

```{r}
site_sf <- read_rds(path_ls$geo_site) %>%
  filter(abbr %in% c("angelo", "carrizo", "elkhorn", "jasper", "mclann", "ucsc")) %>%
  mutate(short_name = case_when(
    abbr == "angelo" ~ "Angelo",
    abbr == "carrizo" ~ "Carrizo Plain",
    abbr == "elkhorn" ~ "Elkhorn Slough",
    abbr == "jasper" ~ "Jasper Ridge",
    abbr == "mclann" ~ "McLaughlin",
    abbr == "ucsc" ~ "UC Santa Cruz"
  ))
site_sf %>%
  as_tibble() %>% # drop geometry
  knitr::kable()
```

Get grassland coverage. Land cover type data on 0.05 degree resolution maunually downloaded from https://lpdaac.usgs.gov/products/mcd12c1v006/.

```{r}
cfp_sf <- st_read(path_ls$geo_cfp) %>%
  filter(
    NAME == "California Floristic Province",
    Type == "hotspot area"
  )

sds <- terra::sds(paste0(path_ls$geo_grass, "MCD12C1.A2020001.006.2021362215328.hdf"))
lc_ras <- raster::raster(sds[1])
raster::extent(lc_ras) <- c(-180, 180, -90, 90)
raster::projection(lc_ras) <- sp::CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
lc_ras <- raster::flip(lc_ras)
lc_ras <- raster::crop(lc_ras, raster::extent(cfp_sf))
lc_ras <- raster::mask(lc_ras, cfp_sf)

grass_ras <- lc_ras
grass_ras[grass_ras != 10] <- NA # User guide at https://lpdaac.usgs.gov/documents/101/MCD12_User_Guide_V6.pdf
grass_ras[grass_ras == 10] <- 1
grass_df <- raster::as.data.frame(grass_ras, xy = T) %>%
  drop_na()
```

```{r fig.cap="Map of experimental and observational study sites"}
set.seed(618)
ggplot() +
  geom_sf(
    data = rnaturalearth::ne_states(
      country = c("Mexico", "United States of America"),
      returnclass = "sf"
    ),
    fill = NA,
    color = alpha("black", .1)
  ) +
  geom_sf(
    data = cfp_sf,
    color = alpha("black", .3),
    fill = alpha("white", .1)
  ) +
  geom_raster(
    data = grass_df,
    aes(x = x, y = y),
    fill = "yellow green"
  ) +
  geom_sf(data = site_sf, color = "red") +
  ggrepel::geom_label_repel(
    data = site_sf,
    mapping = aes(label = short_name, geometry = geometry),
    stat = "sf_coordinates",
    size = 3,
    color = "red",
    fill = NA,
    min.segment.length = 0,
    label.padding = unit(.1, "lines"),
    label.size = NA
  ) +
  labs(x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(-126, -114), ylim = c(28, 44)) +
  scale_x_continuous(breaks = c(-125, -120, -115)) +
  scale_y_continuous(breaks = c(30, 35, 40))
```

```{r}
rm(list = setdiff(ls(), "path_ls"))
```

### Site climate change

TODO: get PRISM/CHELSA data on six sites and plot climate change; may compare with CTI and CPI changes
