---
title: "Main figures"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/utils/setup.R")
```

# Species climate niches

```{r}
source("scripts/plot-main-niche.R")
```

We estimated species' climate niches from the GBIF occurrence data and CHELSA climate data in the California Floristic Province (CFP). In brief, we extracted the mean annual temperature and precipitation for each of the `r n_spp` species with more than 100 occurrences at each of the `r n_occ` locations. For each species, we then calculated the average (median) temperature and precipitation across all locations as its climate niche centroids. We visualized all the species in both the geographical space (longitude and latitude) and climate space (temperature and precipitation).

For illustration, we chose two example species: *Danthonia californica* and *Stipa pulchra*. Both are important native species in California and abundantly distributed. *D. californica* (`r n_occ_cool` occurrences) is a northern species and occurs in cool and wet climates, while *S. pulchra* (`r n_occ_warm` occurrences) is a southern species and occurs in warm and dry climates. We visualized the two example species in the context of all the species.

Species distributions show clear patterns in both geographical space and climate space.

```{r niche-fig, fig.width=7.5, fig.height=7.5*1.168}
#| fig.cap = "Grassland species occur along geographic and climatic gradients to estimate ecological niches.
#|  (A) Occurrence of grassland species in California Floristic Province, retrieved from the GBIF data (gray points), highlighting two example species *Danthonia californica* in the north (blue points) and *Stipa pulchra* in the south (red points).
#|  (B) Species occurrence in climate space of mean annual temperature (°C) and mean annual precipitation (mm), retrieved from the CHELSA data (gray), highlighting the same two example species *Danthonia californica* under cool and wet climates (blue) and *Stipa pulchra* under warm and dry climates (red).
#|  (C) Ecological niche estimates from average temperature and precipitation for each species (median ± standard error, gray), highlighting the two species (blue and red)."
niche_gg
rm(list = ls())
```

# Community climate shifts

Let us begin with the theory of moment method estimation, which is a generalization of the CTI/CPI approach. For location $i$, species $j$, denote species abundance (point intercept or percent cover) as $A_{ij}$, and species trait (e.g., optimal temperature) as $T_j$. At the community level, the expected trait (i.e., first moment), which is equivalent to CTI/CPI, is

$$
\mathrm{E}(T_i) = \overline{T_i} = \frac{\sum_j A_{ij}T_j}{\sum_j A_{ij}}.
$$

## Experimental community shifts

```{r}
source("scripts/plot-main-exp.R")
```

In the Jasper Ridge Global Change Experiment (JRGCE), treatments consisted of four global change factors---temperature, precipitation, CO~2~, and nitrogen---at either ambient or an elevated level, applied in a complete factorial, randomized block design with eight replicates. Thirty-six circular plots 2 m in diameter, subdivided into four equal-sized subplots, were established in the summer of 1997. The 1997‒1998 growing season (year 1) was a pre-treatment year, after which treatments were applied for 16 consecutive growing seasons (years 2‒17), from the time of germination (November) to plant senescence (June).

Here we focus on the warming treatment, which was applied via heaters suspended above the circular plots. The warming treatment increased from +80 W m^--2^ (years 2‒5), to +100 W m^--2^ (years 6‒12), to +250 W m^--2^ (years 13‒17). In total, there were `r n_plt["ambient"]` ambient subplots and `r n_plt["warming"]` warming subplots. (Explain lumping all non-warming treatments to arrive at large sample sizes.)

The grassland community composition was measured using the point intercept method. (Fill in details.)

In each year, for each subplot $i$, species $j$, we assigned the pin count as species abundance $A_{ij}$ and species climate (temperature and precipitation) niche centroid as $T_j$. The community weighted mean (temperature, CTI, or precipitation, CPI) was calculated using the above equation. We compared the CTI/CPI for ambient vs. warming treatments and tested their significance.

The results show warmed plots have higher CTI and lower CPI compared to ambient plots, suggesting warmed plots are dominated by species likely to occur in warmer and drier climates. The differences become more significant with stronger heating treatments over three phases. Another explanation might be the accumulation effects of the grassland community after receiving long-term warming treatments.

```{r exp-fig, fig.width=10, fig.height=6.18}
#| fig.cap="Grassland community shifts from long-term experiments. Warming treatment causes communities to shift dominance to warmer (CTI, °C) and drier (CPI, mm) species in the Jasper Ridge Global Change Experiment (ambient plots in black, warming plots in red, background orange shades denote the pre-treatment year 1998, phase I heating +80 W m^–2^ in 1999-2002, phase II heating +100 W m^–2^ in 2003-2009, and phase III heating +250 W m^–2^ in 2010-2014, non-parametric two-sample Wilcoxon test, ns: *p* > 0.05, $*$: *p* ≤ 0.05, $**$: *p* ≤ 0.01, $***$: *p* ≤ 0.001)."
exp_gg
rm(list = ls())
```

## Observational community shifts

```{r}
source("scripts/plot-main-obs.R")
```

We collected grassland community composition from 12 sites in the CFP with more than \* years of observations. (Fill in details from Josie's dissertation.)

We applied a similar approach to these long-term observational data. In each year, for each plot $i$, species $j$, we assigned species abundance $A_{ij}$ and species climate (temperature and precipitation) niche centroid as $T_j$. The CTI and CPI were calculated using the same equation. We compared the CTI/CPI over the years and tested their significance.

The results show eight out of the 12 sites have significant increases in CTI and significant decreases in CPI over the years. The remaining four sites do not experience significant changes in CTI and CPI. These changes suggest that the majority of these grassland communities are shifting composition to be dominated by warmer and drier species, which is consistent with climate warming in the region.

```{r obs-fig, fig.width=10, fig.height=10*1.618}
#| fig.cap="Grassland community shifts from long-term observations. Consistent with climate warming and drying, grassland communities shift dominance to warmer (CTI) and drier (CPI) species in 12 long-term observational sites across the California Floristic Province (red trend lines by linear regression, significant slopes by *t*-test, *p* < 0.05)."
obs_gg
rm(list = ls())
```
