---
title: "Reporting statistics"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/_setup.R")
```

# Main text

# Supplementary materials

```{r compare-bioclim}
source("scripts/compare-bioclim.R")
```

Species climate niche can be quantified by variables other than mean annual temperature and precipitation. We downloaded relevant bioclimatic variables defined by [WorldClim](https://www.worldclim.org/data/bioclim.html) and extracted to all the occurrence locations in the CFP. 

Here we show that annual mean temperature (BIO1) used in this study is highly correlated with seasonal temperatures, i.e., mean temperature of wettest quarter (BIO8, *r* = `r round(tmp_cor["bio1", "bio8"], 3)`), mean temperature of driest quarter (BIO9, *r* = `r round(tmp_cor["bio1", "bio9"], 3)`), mean temperature of warmest quarter (BIO10, *r* = `r round(tmp_cor["bio1", "bio10"], 3)`), and mean temperature of coldest quarter (BIO11, *r* = `r round(tmp_cor["bio1", "bio11"], 3)`).

We also show that annual precipitation (BIO12) used in this study is highly correlated with winter precipitations, i.e., precipitation of wettest month (BIO13, *r* = `r round(ppt_cor["bio12", "bio13"], 3)`), precipitation of wettest quarter (BIO16, *r* = `r round(ppt_cor["bio12", "bio16"], 3)`), and precipitation of coldest quarter (BIO19, *r* = `r round(ppt_cor["bio12", "bio19"], 3)`).

Because we are interested in the change of community-weighted mean climate, the high correlations above suggest that our results are robust to the bioclimatic variables chosen. 

# Computing environment

```{r}
sessionInfo()
```

