---
title: "Experimental community shift analyses"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/utils/setup.R")
```

# Methods

Let us begin with the theory of moment method estimation, which is a generalization of the CTI/CPI approach. For location $i$, species $j$, denote species abundance (point intercept or percent cover) as $A_{ij}$, and species trait (e.g., optimal temperature) as $T_j$. At the community level, the expected trait (i.e., first moment), which is equivalent to CTI/CPI, is

$$
\mathrm{E}(T_i) = \bar{T_i} = \frac{\sum_j A_{ij}T_j}{\sum_j A_{ij}}.
$$

Furthermore, the trait variance (i.e., second central moment), using the variance-expectation formula, is

$$
\mathrm{Var}(T_i) = \mathrm{E}[T_i^2] - \mathrm{E}[T_i]^2 = \frac{\sum_j A_{ij}T_j^2}{\sum_j A_{ij}} - \bar{T_i}^2.
$$

A pair of $\{\mathrm{E}(T_i), \mathrm{Var}(T_i)\}$ characterizes the trait values under the distribution of species abundance. Operationally, we might use a pair of mean and standard deviation, $\{ \mu(T_i) = \mathrm{E}(T_i), \sigma(T_i) = \sqrt{\mathrm{Var}(T_i)}\}$, so that they are comparable with the same unit. 

Show available sampling year for plot by site for experimental and observational data.

```{r data-avail}
read_rds(.path$com_exp) %>%
  distinct(site, plot, year) %>%
  ggplot(aes(year, plot)) +
  geom_point() +
  facet_wrap(~site, scales = "free")

read_rds(.path$com_obs) %>%
  distinct(site, plot, year) %>%
  ggplot(aes(year, plot)) +
  geom_point() +
  facet_wrap(~site, scales = "free")
```

Calculate CWM using niche estimates for experimental and observational data.

```{r cal-cwm}
source("scripts/calculate-cwm.R")
```

Load customized plotting functions.

```{r plot-cwm-exp}
source("scripts/plot-cwm-exp.R")
```

# JRGCE 

In the following figures, left-right indicates warming-only vs. warming-all treatments. Warming-only treatments have fewer plots; all warming treatments have more plots, which average over all the other precip, CO2, nitrogen treatments. Top-bottom indicates mean or std dev of the community indexes.

## CTI ~ warming treatment

```{r jrgce-cti-warming, fig.cap="Community Temperature Index (CTI) statistics respond to warming treatments}
plot_cwm_jrgce(cwm_tag = "mean_cti", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cti", trt_tag = "tmp_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cti", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cti", trt_tag = "tmp_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

## CPI ~ warming treatment

```{r jrgce-cpi-warming, fig.cap="Community Precipitation Index (CPI) statistics respond to warming treatments}
plot_cwm_jrgce(cwm_tag = "mean_cpi", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cpi", trt_tag = "tmp_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cpi", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cpi", trt_tag = "tmp_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

## CVI ~ warming treatment

```{r jrgce-cvi-warming, fig.cap="Community VPD Index (CVI) statistics respond to warming treatments}
plot_cwm_jrgce(cwm_tag = "mean_cvi", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cvi", trt_tag = "tmp_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cvi", trt_tag = "tmp_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cvi", trt_tag = "tmp_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

## CTI ~ watering treatment

```{r jrgce-cti-watering, fig.cap="Community Temperature Index (CTI) statistics respond to watering treatments}
plot_cwm_jrgce(cwm_tag = "mean_cti", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cti", trt_tag = "ppt_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cti", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cti", trt_tag = "ppt_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

## CPI ~ watering treatment

```{r jrgce-cpi-watering, fig.cap="Community Precipitation Index (CPI) statistics respond to watering treatments}
plot_cwm_jrgce(cwm_tag = "mean_cpi", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cpi", trt_tag = "ppt_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cpi", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cpi", trt_tag = "ppt_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

## CVI ~ watering treatment

```{r jrgce-cvi-watering, fig.cap="Community VPD Index (CVI) statistics respond to warming treatments}
plot_cwm_jrgce(cwm_tag = "mean_cvi", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "mean_cvi", trt_tag = "ppt_all") +
  plot_cwm_jrgce(cwm_tag = "sd_cvi", trt_tag = "ppt_only") +
  plot_cwm_jrgce(cwm_tag = "sd_cvi", trt_tag = "ppt_all") +
  plot_annotation(tag_levels = "A") +
  plot_layout(2, 2)
```

# McLaughlin Water Experiment

In the following figures, left-right indicates non-serpentine vs. serpentine soil types. Top-bottom indicates mean or std dev of the community indexes.

## CTI ~ watering treatment

```{r mclexp-cti-watering, fig.cap="Community Temperature Index (CTI) statistics respond to watering treatments}
plot_cwm_mclexp(cwm_tag = "mean_cti", trt_tag = "watering") +
  plot_cwm_mclexp(cwm_tag = "sd_cti", trt_tag = "watering") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CPI ~ watering treatment

```{r mclexp-cpi-watering, fig.cap="Community Precipitation Index (CPI) statistics respond to watering treatments}
plot_cwm_mclexp(cwm_tag = "mean_cpi", trt_tag = "watering") +
  plot_cwm_mclexp(cwm_tag = "sd_cpi", trt_tag = "watering") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CVI ~ watering treatment

```{r mclexp-cvi-watering, fig.cap="Community VPD Index (CVI) statistics respond to watering treatments}
plot_cwm_mclexp(cwm_tag = "mean_cvi", trt_tag = "watering") +
  plot_cwm_mclexp(cwm_tag = "sd_cvi", trt_tag = "watering") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CTI ~ drought treatment

```{r mclexp-cti-drought, fig.cap="Community Temperature Index (CTI) statistics respond to drought treatments}
plot_cwm_mclexp(cwm_tag = "mean_cti", trt_tag = "drought") +
  plot_cwm_mclexp(cwm_tag = "sd_cti", trt_tag = "drought") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CPI ~ drought treatment

```{r mclexp-cpi-drought, fig.cap="Community Precipitation Index (CPI) statistics respond to drought treatments}
plot_cwm_mclexp(cwm_tag = "mean_cpi", trt_tag = "drought") +
  plot_cwm_mclexp(cwm_tag = "sd_cpi", trt_tag = "drought") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CVI ~ drought treatment

```{r mclexp-cvi-drought, fig.cap="Community VPD Index (CVI) statistics respond to drought treatments}
plot_cwm_mclexp(cwm_tag = "mean_cvi", trt_tag = "drought") +
  plot_cwm_mclexp(cwm_tag = "sd_cvi", trt_tag = "drought") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

# Santa Cruz IDE

Michael said Marshall Field has lots of native species, while Arboretum and YLR are highly invaded (2/16/22). That might explain the difference among those sites. 

In the following figures, left-right indicates different sites; top-bottom indicates mean or std dev of the community indexes.

## CTI ~ drought treatment

```{r scide-cti-drought, fig.cap="Community Temperature Index (CTI) statistics respond to drought treatments"}
plot_cwm_scide(cwm_tag = "mean_cti") +
  plot_cwm_scide(cwm_tag = "sd_cti") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CPI ~ drought treatment

```{r scide-cpi-drought, fig.cap="Community Precipitation Index (CPI) statistics respond to drought treatments"}
plot_cwm_scide(cwm_tag = "mean_cpi") +
  plot_cwm_scide(cwm_tag = "sd_cpi") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

## CVI ~ drought treatment
```{r scide-cvi-drought, fig.cap="Community VPD Index (CVI) statistics respond to drought treatments"}
plot_cwm_scide(cwm_tag = "mean_cvi") +
  plot_cwm_scide(cwm_tag = "sd_cvi") +
  plot_annotation(tag_levels = "A") +
  plot_layout(1, 2)
```

```{r clean}
rm(list = ls())
```
