---
title: "Observational community shift analyses"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
    fig_width: 15
    fig_height: 15 
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
    fig_width: 15
    fig_height: 15 
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/utils/setup.R")
```

# Community change

Show available sampling year for plot by site for observational data.

```{r data-avail}
read_rds(.path$com_obs) %>%
  distinct(site, plot, year) %>%
  ggplot(aes(year, plot)) +
  geom_point() +
  facet_wrap(~site, scales = "free")
```

Calculate CWM using niche estimates for experimental and observational data.

```{r cal-cwm}
source("scripts/calculate-cwm.R")
```

Load customized plotting functions.

```{r plot-cwm-obs}
source("scripts/plot-cwm-obs.R")
```

## CTI ~ year

Community temperature niches change over years.

CTI community-level mean in several sites increases over years, suggesting communities are becoming "warmer," i.e., thermophilization.

```{r obs-mean-cti, fig.cap="Compare mean(CTI) across sites over years"}
plot_cwm_obs("mean_cti")
```

CTI community-level standard deviation.

```{r obs-sd-cti, fig.cap="Compare sd(CTI) across sites over years"}
plot_cwm_obs("sd_cti")
```

## CPI ~ year

Community precipitation niches change over years.

CPI community-level mean in several sites decreases over years, suggesting communities are becoming "drier," i.e., mesophilization.

```{r obs-mean-cpi, fig.cap="Compare mean(CPI) across sites over years"}
plot_cwm_obs("mean_cpi")
```

CPI community-level standard deviation.

```{r obs-sd-cpi, fig.cap="Compare sd(CPI) across sites over years"}
plot_cwm_obs("sd_cpi")
```

## CVI ~ year

Community VPDmax niches change over years.

CVI community-level mean in several sites changes over years, but no clear direction.

```{r obs-mean-cvi, fig.cap="Compare mean(CVI) across sites over years"}
plot_cwm_obs("mean_cvi")
```

CVI community-level standard deviation.

```{r obs-sd-cvi, fig.cap="Compare sd(CVI) across sites over years"}
plot_cwm_obs("sd_cvi")
```


<!-- ``` -->
