---
title: "Observational community shift analyses"
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
output:
  bookdown::html_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
    fig_width: 15
    fig_height: 15 
  bookdown::word_document2: 
    fig_caption: true
    number_sections: true
    global_numbering: true
    fig_width: 15
    fig_height: 15 
---

```{r setup, echo=FALSE, message=FALSE}
source("scripts/utils/setup.R")
```

# Community change

Show available sampling year for plot by site for observational data.

```{r data-avail}
read_rds(.path$com_obs) %>%
  distinct(site, plot, year) %>%
  ggplot(aes(year, plot)) +
  geom_point() +
  facet_wrap(~site, scales = "free")
```

Calculate CWM using niche estimates for experimental and observational data.

```{r cal-cwm}
source("scripts/calculate-cwm.R")
```

Load customized plotting functions.

```{r plot-cwm-obs}
source("scripts/plot-cwm-obs.R")
```

## CTI ~ year

Community temperature niches change over years.

CTI community-level mean in several sites increases over years, suggesting communities are becoming "warmer," i.e., thermophilization.

```{r obs-mean-cti, fig.cap="Compare mean(CTI) across sites over years"}
plot_cwm_obs("mean_cti")
```

CTI community-level standard deviation.

```{r obs-sd-cti, fig.cap="Compare sd(CTI) across sites over years"}
plot_cwm_obs("sd_cti")
```

## CPI ~ year

Community precipitation niches change over years.

CPI community-level mean in several sites decreases over years, suggesting communities are becoming "drier," i.e., mesophilization.

```{r obs-mean-cpi, fig.cap="Compare mean(CPI) across sites over years"}
plot_cwm_obs("mean_cpi")
```

CPI community-level standard deviation.

```{r obs-sd-cpi, fig.cap="Compare sd(CPI) across sites over years"}
plot_cwm_obs("sd_cpi")
```

## CVI ~ year

Community VPDmax niches change over years.

CVI community-level mean in several sites changes over years, but no clear direction.

```{r obs-mean-cvi, fig.cap="Compare mean(CVI) across sites over years"}
plot_cwm_obs("mean_cvi")
```

CVI community-level standard deviation.

```{r obs-sd-cvi, fig.cap="Compare sd(CVI) across sites over years"}
plot_cwm_obs("sd_cvi")
```

# Climate change

Observational site climate change over the study period.

```{r}
site_tbl <- read_rds("data/community/site-info.rds")
```

Download PRISM data (to be updated with Chelsa). Set PRISM dir first.

```{r}
prism::prism_set_dl_dir("data/climate/prism/")
```

Download (not run, only once).

```{r, eval=FALSE}
range(exp_tbl$year) # 1998:2021
range(obs_tbl$year) # 1983:2019
prism::get_prism_annual("tmean", years = 1983:2021, keepZip = FALSE)
prism::get_prism_annual("ppt", years = 1983:2021, keepZip = FALSE)
prism::get_prism_annual("vpdmax", years = 1983:2021, keepZip = FALSE)
```

Extract site climate from PRISM.

```{r}
site_tmp_tbl <- prism::prism_archive_subset(
  "tmean", "annual",
  years = 1983:2020
) %>%
  prism::pd_stack() %>%
  raster::extract(site_tbl) %>%
  bind_cols(site = site_tbl$abbr, .) %>%
  pivot_longer(-site, names_to = "year", values_to = "tmp") %>%
  mutate(year = year %>%
    str_sub(start = -8L, end = -5L) %>%
    as.integer())

site_ppt_tbl <- prism::prism_archive_subset(
  "ppt", "annual",
  years = 1983:2020
) %>%
  prism::pd_stack() %>%
  raster::extract(site_tbl) %>%
  bind_cols(site = site_tbl$abbr, .) %>%
  pivot_longer(-site, names_to = "year", values_to = "ppt") %>%
  mutate(year = year %>%
    str_sub(start = -8L, end = -5L) %>%
    as.integer())

site_vpd_tbl <- prism::prism_archive_subset(
  "vpdmax", "annual",
  years = 1983:2020
) %>%
  prism::pd_stack() %>%
  raster::extract(site_tbl) %>%
  bind_cols(site = site_tbl$abbr, .) %>%
  pivot_longer(-site, names_to = "year", values_to = "vpd") %>%
  mutate(year = year %>%
    str_sub(start = -8L, end = -5L) %>%
    as.integer())
```

Assemble site climate data.

```{r}
site_clim_tbl <- site_tmp_tbl %>%
  full_join(site_ppt_tbl, by = c("site", "year")) %>%
  full_join(site_vpd_tbl, by = c("site", "year")) %>%
  right_join(obs_tbl %>%
    distinct(site, year), by = c("site", "year")) %>%
  arrange(site, year)
```

Plot temperature change.

```{r, fig.cap="Observational site temperature change across years"}
site_clim_tbl %>%
  ggplot(aes(year, tmp)) +
  geom_point() +
  geom_smooth(
    method = "lm", se = FALSE,
    color = "red", lty = "dashed"
  ) +
  ggpubr::stat_cor(aes(label = ..p.label..),
    p.accuracy = 0.05,
    color = "red"
  ) +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Mean annual temperature (°C)"
  )
```

Plot precipitation change.

```{r, fig.cap="Observational site precipitation change across years"}
site_clim_tbl %>%
  ggplot(aes(year, ppt)) +
  geom_point() +
  geom_smooth(
    method = "lm", se = FALSE,
    color = "red", lty = "dashed"
  ) +
  ggpubr::stat_cor(aes(label = ..p.label..),
    p.accuracy = 0.05,
    color = "red"
  ) +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Mean annual precipitation (mm)"
  )
```

Plot VPDmax change.

```{r, fig.cap="Observational site VPDmax change across years"}
site_clim_tbl %>%
  ggplot(aes(year, vpd)) +
  geom_point() +
  geom_smooth(
    method = "lm", se = FALSE,
    color = "red", lty = "dashed"
  ) +
  ggpubr::stat_cor(aes(label = ..p.label..),
    p.accuracy = 0.05,
    color = "red"
  ) +
  facet_wrap(~site, scales = "free_y") +
  labs(
    x = "Year", y = "Mean VPDmax (hPa)"
  )
```

Remove intermediate objects.

```{r}
rm(site_clim_tbl, site_ppt_tbl, site_tbl, site_tmp_tbl, site_vpd_tbl)
rm(exp_file, exp_tbl, niche_file, niche_tbl, obs_file, obs_tbl)
```
