---
title: "Replicate key results"
author: "Kai Zhu"
date: "1/21/2022"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

# Objectives

1. Replicate Josie's key results on CTI change
1. Compare with site-level climate change

# Community climate index

Modified from Josie's code `Com_MajorAnalyses.Rmd` section `# Changes in wCTI over time`.

```{r}
cci_dat <- read_csv("Analysis/CTI Data/AllSite_CTI_08_2020.csv") %>%
  select(-`...1`) %>%
  filter(
    species.name != "Linanthus jepsonii",
    species.name != "Linanthus latisectus",
    relcov >= 0
  )

cci_plot_sum <- cci_dat %>%
  group_by(site, plot, year) %>%
  mutate(
    relCTI = relcov * meanT,
    relCPI = relcov * meanp
  ) %>%
  summarise(
    plotCTI = sum(relCTI),
    plotCPI = sum(relCPI)
  ) %>%
  arrange(site, plot, year)

cci_site_sum <- cci_plot_sum %>% # from plot summary, can redo using indiviual obs
  group_by(site, year) %>%
  summarise(
    meanCTI = mean(plotCTI),
    seCTI = plotrix::std.error(plotCTI),
    meanCPI = mean(plotCPI),
    seCPI = plotrix::std.error(plotCPI)
  )

cci_site_gg <- ggplot(cci_site_sum, aes(
  x = year,
  y = meanCTI, ymin = meanCTI - seCTI, ymax = meanCTI + seCTI
)) +
  geom_pointrange() +
  geom_line() +
  facet_wrap(~site, scales = "free", nrow = 2)

cci_site_gg +
  labs(
    title = "Replicate Josie's Figure CTI",
    x = "Year",
    y = "Community Temperature Index (CTI, degC)"
  )
```

Note that the goal was to replicate Josie's manuscript Figure CTI. The data look right, but sites seem to be mislabelled. 

# Site climate data

Modified from Josie's code `Com_MajorAnalyses.Rmd` subsection `## CTI vs actual temp`.

```{r}
clim_dat <- read_csv("Analysis/ClimData/PRISM_wtryr_alldata.csv") %>%
  select(-`...1`)

clim_site_sum <- cci_dat %>%
  distinct(site, year) %>%
  left_join(clim_dat, by = c("site" = "site", "year" = "wateryear")) %>%
  mutate(grow_meanT = (grow_minT + grow_maxT) / 2) # grow_meanT is highly correlated with mean_meanT

clim_site_gg <- clim_site_sum %>%
  ggplot(aes(x = year, y = grow_meanT)) + # ymin = grow_minT, ymax = grow_maxT
  geom_point() +
  geom_line() +
  facet_wrap(~site, scales = "free", nrow = 2)

clim_site_gg +
  labs(
    title = "Replicate Josie's Figure weather",
    x = "Year",
    y = "Growing season temperature (degC)"
  )
```

Note that the goal was to replicate Josie's manuscript Figure weather and Table weather models. It looks right to me (no mislabel).

# Combine climate change and community change

Time-series plots of site-level climate change and site-level community climate index change.

```{r fig.height = 10}
(clim_site_gg +
  geom_smooth(method = "lm") +
  labs(title = "(a) Climate change", x = "", y = "Growing season temperature (degC)")) /
  (cci_site_gg +
    geom_smooth(method = "lm") +
    labs(title = "(b) Community change", x = "Year", y = "Community Temperature Index (CTI, degC)"))
```

I think the comparison looks very interesting. Climate change seems to drive community composition change.

Directly compare site-level climate values to community climate index seems less intuitive. 

```{r}
clim_site_sum %>%
  full_join(cci_site_sum, by = c("site", "year")) %>%
  ggplot(aes(
    x = grow_meanT,
    y = meanCTI, ymin = meanCTI - seCTI, ymax = meanCTI + seCTI
  )) +
  geom_pointrange() +
  facet_wrap(~site, scales = "free", nrow = 2) +
  geom_smooth(method = "lm") +
  labs(
    x = "Growing season temperature (degC)",
    y = "Community Temperature Index (CTI, degC)"
  )
```
