---
title: "5 Community shift"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5 Community shift}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = here::here()) # knit from project directory
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  cache = T,
  comment = "#>",
  fig.width = 8,
  fig.height = 8
)
```

```{r setup}
library(grassland)
theme_set(ggthemes::theme_few())
```

Plot guild percentage change over time from observational and experimental data.

```{r}
gg_guild_percentage_obs <- plot_guild_percentage(option = "obs", dat_community = dat_community)
gg_guild_percentage_obs
gg_guild_percentage_exp <- plot_guild_percentage(option = "exp", dat_community = dat_community)
gg_guild_percentage_exp
```

Statistically test species climate trait (temperature or precipitation) change using weighted LM or LME.

```{r}
df_trait_change_obs <- test_trait_change_all(dat_community, dat_niche, option = "obs")
df_trait_change_obs %>% knitr::kable()
df_trait_change_exp <- test_trait_change_all(dat_community, dat_niche, option = "exp")
df_trait_change_exp %>% knitr::kable()
df_summ_trait_change_obs <- summ_change(df_trait_change_obs, option = "obs")
```

Overall, the 12 observational sites had significant increases in temperature trait (`r df_trait_change_obs %>% filter(site == "all") %>% filter(trait == "tmp") %>% pull(estimate) %>% signif(3)` °C yr^-1^) and significant decreases in precipitation trait (`r df_trait_change_obs %>% filter(site == "all") %>% filter(trait == "ppt") %>% pull(estimate) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^) over the years. 

By site, `r df_summ_trait_change_obs %>% pull(n_sig) %>% min()` out of the 12 sites had significant increases in temperature trait and significant decreases in precipitation trait over the years. Among all 12 sites, the temperature traits increases at a rate ranging from `r df_summ_trait_change_obs %>% filter(trait == "tmp") %>% pull(min) %>% signif(3)` °C yr^-1^ to `r df_summ_trait_change_obs %>% filter(trait == "tmp") %>% pull(max) %>% signif(3)` °C yr^-1^; the precipitation trait decreases at a rate ranging from `r df_summ_trait_change_obs %>% filter(trait == "ppt") %>% pull(max) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^ to `r df_summ_trait_change_obs %>% filter(trait == "ppt") %>% pull(min) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^.

Comparing warming plots to ambient plots, pooling data from all years in each phase, no significant difference was detected in Phase I or II, whereas Phase III exhibited a significant increase in temperature trait of `r df_trait_change_exp %>% filter(exp == "jrgce", grp == "Phase III") %>% filter(trait == "tmp") %>% pull(estimate) %>% signif(3)` °C and a significant decrease in precipitation trait of `r df_trait_change_exp %>% filter(exp == "jrgce", grp == "Phase III") %>% filter(trait == "ppt") %>% pull(estimate) %>%magrittr::multiply_by(-1)` mm.

Plot the above comparisons and tests.

```{r}
gg_trait_change_obs <- plot_trait_change(option = "obs", dat_community, dat_niche)
gg_trait_change_obs
gg_trait_change_exp <- plot_trait_change(option = "exp", dat_community, dat_niche)
gg_trait_change_exp
```

Calculate Community Temperature Index (CTI) and Community Precipitation Index (CPI) in observations and experiments.

```{r, eval = F}
dat_index <- calc_community_index(dat_niche = dat_niche, dat_community = dat_community)
```

Statistically test CTI and CPI changes using LM or LME.

```{r}
df_index_change_obs <- test_index_change_all(dat_index, option = "obs")
df_index_change_obs %>% knitr::kable()
df_index_change_exp <- test_index_change_all(dat_index, option = "exp")
df_index_change_exp %>% knitr::kable()
df_summ_index_change_obs <- summ_change(df_index_change_obs, option = "obs")
```

Overall, the 12 observational sites had significant thermophilization (increases in CTI) (`r df_index_change_obs %>% filter(site == "all") %>% filter(index == "cti") %>% pull(estimate) %>% signif(3)` °C yr^-1^) and significant xerophilization (decreases in CPI) (`r df_index_change_obs %>% filter(site == "all") %>% filter(index == "cpi") %>% pull(estimate) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^) over the years. 

By site, `r df_summ_index_change_obs %>% pull(n_sig) %>% min()` out of the 12 sites had significant increases in CTI and significant decreases in CPI over the years. Among all 12 sites, the mean thermophilization rate (increasing CTI per year) ranged from `r df_summ_index_change_obs %>% filter(index == "cti") %>% pull(min) %>% signif(3)` °C yr^-1^ to `r df_summ_index_change_obs %>% filter(index == "cti") %>% pull(max) %>% signif(3)` °C yr^-1^; the mean xerophilization rate (decreasing CPI per year) ranged from `r df_summ_index_change_obs %>% filter(index == "cpi") %>% pull(max) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^ to `r df_summ_index_change_obs %>% filter(index == "cpi") %>% pull(min) %>% signif(3) %>% magrittr::multiply_by(-1)` mm yr^-1^.

Comparing warming plots to ambient plots, pooling data from all years in each phase, no significant difference was detected in Phase I or II, whereas Phase III exhibited a significant thermophilization (increase in CTI) of `r df_index_change_exp %>% filter(exp == "jrgce", grp == "Phase III") %>% filter(index == "cti") %>% pull(estimate) %>% signif(3)` °C and a significant decrease in precipitation trait of `r df_index_change_exp %>% filter(exp == "jrgce", grp == "Phase III") %>% filter(index == "cpi") %>% pull(estimate) %>% signif(3) %>% magrittr::multiply_by(-1)` mm.

Plot the above comparisons and tests.

```{r}
gg_community_index_obs <- plot_community_index(option = "obs", dat_index = dat_index)
gg_community_index_obs
gg_community_index_exp <- plot_community_index(option = "exp", dat_index = dat_index, experiment = "jrgce", treatment = "warming")
gg_community_index_exp
```

Due to the high correlation of species temperature and precipitation niches, CTI and CPI shifts at the community level appear highly coupled. Synthesize the coupled CTI-CPI shifts, both visually and numerically.

```{r, eval = F}
dat_shift <- calc_community_shift(dat_index = dat_index)
```

```{r}
gg_community_shift <- plot_community_shift(dat_shift, dat_niche)
gg_community_shift$combined
```

```{r}
df_summ_shift_coupling <- summ_shift_coupling(dat_shift)
df_summ_shift_coupling
```

A community thermophilization of 0.1 °C corresponded to a xerophilization of `r df_summ_shift_coupling %>% filter(ratio == "cpi_to_cti") %>% filter(group == "Observation") %>% pull(median) %>% signif(3) %>% abs()` mm from the observations and `r df_summ_shift_coupling %>% filter(ratio == "cpi_to_cti") %>% filter(group == "Experiment") %>% pull(median) %>% signif(3) %>% abs()` mm from the experiment. Likewise, a community xerophilization of 10 mm corresponded to a thermophilization of `r df_summ_shift_coupling %>% filter(ratio == "cti_to_cpi") %>% filter(group == "Observation") %>% pull(median) %>% signif(3) %>% abs()` °C from the observations and `r df_summ_shift_coupling %>% filter(ratio == "cti_to_cpi") %>% filter(group == "Experiment") %>% pull(median) %>% signif(3) %>% abs()` °C from the experiment.

Output figures and R data files.

```{r, eval = F}
save_figure(out = list(
  guild_perc_obs = gg_guild_percentage_obs,
  guild_perc_exp = gg_guild_percentage_exp,
  trait_change_obs = gg_trait_change_obs,
  trait_change_exp = gg_trait_change_exp,
  community_index_obs = gg_community_index_obs,
  community_index_exp = gg_community_index_exp,
  community_shift = gg_community_shift$combined
))
save_rda(out = list(dat_index, dat_shift))
```
