---
title: "6 Other experiments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6 Other experiments}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = here::here()) # knit from project directory
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  cache = T,
  comment = "#>",
  fig.width = 8,
  fig.height = 8,
  dpi = 300
)
```

```{r setup, cache = F, message = F}
library(grassland)
theme_set(ggthemes::theme_few())
```

## Jasper Ridge Global Change Experiment watering treatment

```{r}
(gg_community_index_exp_jrgce_water <- plot_community_index(option = "exp", dat_index, experiment = "jrgce", treatment = "watering"))
```
Watering treatment in the Jasper Ridge Global Change Experiment generally decreased CTI (째C) and increased CPI (mm) over all years, although effects were not significant in most individual years.

## McLaughlin Water Experiment.

```{r}
(gg_community_index_exp_mclexp <- plot_community_index(option = "exp", dat_index, experiment = "mclexp"))
```
Watering treatment in the McLaughlin Water Experiment generally decreased CTI (째C) and increased CPI (mm) in both (A) serpentine and (B) non-serpentine sites over all years, while drought treatments generally increased CTI (째C) and decreased CPI (mm) in (C) the serpentine site over all years. Effects were not significant in most individual years.

## Santa Cruz International Drought Experiment.

```{r}
(gg_community_index_exp_scide <- plot_community_index(option = "exp", dat_index, experiment = "scide"))
```
Drought treatment at three sites, (A) Arboretum, (B) Marshall Field, and (C) Younger Lagoon, in the Santa Cruz International Drought Experiment generally increased CTI (째C) and decreased CPI (mm) over all years. Effects were not significant in most individual years.

```{r}
df_index_change_exp <- test_index_change_all(dat_index, option = "exp", bonferroni = F)
(gg_effect_size <- plot_effect_size(df_index_change_exp))
```


Output figures and R data files.

```{r, eval = F}
save_figure(out = list(
  community_index_exp_jrgce_water = gg_community_index_exp_jrgce_water,
  community_index_exp_mclexp = gg_community_index_exp_mclexp,
  community_index_exp_scide = gg_community_index_exp_scide,
  effect_size = gg_effect_size
))
```
