---
title: "Richness_Analysis"
author: "Josie Lesage"
date: "1/23/2020"
output: html_document
---

# Setup and data import
```{r setup, include=FALSE}
## community analysis
library(codyn)
library(MASS)
library(lme4)
library(nlme)


## misc/both/other
library(oz)
library(plotrix)
library(tidyverse)



knitr::opts_chunk$set(echo = TRUE)
```

```{r com data import, echo=FALSE}
elkhorn_data <- read.csv("Analysis/ComData/elkhorn_data.csv")
elkhorn_data <- elkhorn_data %>%
  select(-X) %>%
  filter(species.name != "Moss",
         species.name != "Thatch",
         species.name != "Bare ground") %>%
  mutate(site = "Elkhorn")
elkhorn_data_rel <- elkhorn_data %>%
  select(year, rep, species.name, intercepts) %>%
  group_by(rep, year) %>%
  mutate(sumintercept = sum(intercepts),
         relcov = intercepts/sumintercept)

mcLserp_data <- read.csv("mclserp_data.csv") %>%
  select(-X) %>%
   rename(species.name = Species_Name)
mcLserp_data <- mcLserp_data %>%
  filter(species.name != "Bare",
         species.name != "Rock",
         species.name != "Unknown forb") %>%
  mutate(site = "McL Serp",
         year = Year)
mcLserp_byquad <- mcLserp_data %>%
  group_by(Site, Year, plot, species.name) %>%
  summarise(Cover = sum(Cover)) %>%
  ungroup() %>%
  group_by(plot, Year) %>%
  mutate(sumcov = sum(Cover),
         relcov = Cover/sumcov)

jasper_data <- read.csv("jasper_data.csv")
jasper_data <- jasper_data %>%
  select(-X) %>%
  filter(cover > 0) %>%
  group_by(uniqueID, year) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>%
  mutate(site = "Jasper")

swan_data <- read.csv("swanton_data.csv")
swan_data <- swan_data %>%
  select(-X) %>%
  filter(species.name != "Moss",
         species.name != "Thatch",
         species.name != "Bare ground") %>%
  mutate(site = "Swanton")
swan_data_rel <- swan_data %>%
  select(year, rep, species.name, intercepts) %>%
  group_by(rep, year) %>%
  mutate(sumintercept = sum(intercepts),
         relcov = intercepts/sumintercept)

angelo_data <- read.csv("angelo_data.csv") %>%
  select(-X)  %>%
  mutate(site = "Angelo")
ang_data_rel <- angelo_data %>%
  select(year, plot, species.name, cover) %>%
  group_by(plot, year) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov)

McLAnn_data <- read.csv("mclann_data.csv") %>%
   rename(species.name = Species_Name)
McLAnn_data <- McLAnn_data %>%
  select(-X) %>%
  filter(species.name != "Bare",
         species.name != "Rock",
         species.name != "Unknown grass",
         species.name != "Unknown forb") %>%
  mutate(site = "McL Ann",
         year = Year)
McLAnn_byquad <- McLAnn_data %>%
  group_by(Site, Year, plot, species.name) %>%
  summarise(Cover = sum(Cover)) %>%
  ungroup() %>%
  group_by(plot, Year) %>%
  mutate(sumcov = sum(Cover),
         relcov = Cover/sumcov)

car_data <- read.csv("carrizo_data.csv") %>%
    rename(species.name = Species.Name)
car_data <- car_data %>%
  select(-X) %>%
  filter(species.name != "litter",
         species.name != "bare",
         species.name != "hole",
         species.name != "moss") %>%
  mutate(site = "Carrizo")
car_data_rel <- car_data %>%
  select(year, Plot, species.name, Intercepts) %>%
  distinct(year, Plot, species.name, .keep_all = TRUE) %>%
  group_by(Plot, year) %>%
  mutate(sumintercept = sum(Intercepts),
         relcov = Intercepts/sumintercept)

ucsc_data <- read.csv("ucsc_data.csv")
ucsc_data <- ucsc_data %>%
  select(-X) %>%
  filter(species.name != "Thatch",
         species.name != "Bare ground") %>%
  mutate(site = "UCSC")
ucsc_data_rel <- ucsc_data %>%
  select(year, rep, species.name, intercepts) %>%
  group_by(rep, year) %>%
  mutate(sumintercept = sum(intercepts),
         relcov = intercepts/sumintercept)


```

```{r prism data import and prep, echo = FALSE}
weather <- read.csv("PRISM_data.csv") %>%
  select(-X)
wtryr_weather <- read.csv("PRISM_wtryr_data.csv") %>%
  select(-X)
```

# Changes in guild-level richness over time
## Pure richness

```{r converting species level cover to guild level cover}
ang_grich <- angelo_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(plot, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAG = 0) %>% add_column(NAGr = 0) %>% add_column(EPG = 0) %>%
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "Angelo")

car_grich <- car_data %>%
  group_by(year, Plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, Plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(Plot, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(EPG = 0) %>% add_column(NPGr = 0) %>%
  add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "Carrizo")

elk_grich <- elkhorn_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(rep, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "Elkhorn")

jasp_grich <- jasper_data %>%
  group_by(year, uniqueID, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, uniqueID) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(uniqueID, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>% 
  add_column(NAGr = 0) %>%  add_column(EAF = 0) %>% add_column(EPF = 0) %>%
  add_column(EPG = 0) %>%  add_column(NPGr = 0) %>% add_column(NPS = 0) %>%
  add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "Jasper")

McLAnn_grich <- McLAnn_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(plot, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "McL Ann")

mcLserp_grich <- mcLserp_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(plot, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "McL Serp")

swan_grich <- swan_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(rep, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(EPG = 0) %>% 
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "Swanton")

ucsc_grich <- ucsc_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(rep, year, rich, guild)  %>%
  spread(guild, rich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', "all", key = "guild", value = "rich") %>%
  group_by(year, guild) %>%
  summarise(meanrich = mean(rich),
            serich = (std.error(rich))) %>%
  mutate(site = "UCSC")

all.grich <- full_join(mcLserp_grich, swan_grich)
all.grich <- full_join(all.grich, elk_grich)
all.grich <- full_join(all.grich, jasp_grich)
all.grich <- full_join(all.grich, ang_grich)
all.grich <- full_join(all.grich, McLAnn_grich)
all.grich <- full_join(all.grich, car_grich)
all.grich <- full_join(all.grich, ucsc_grich)

all.grich <- left_join(all.grich, wtryr_weather, by = c("year", "site")) 

all.grich.means <- all.grich %>%
  group_by(site, guild, year) %>%
  summarise(serich = std.error(serich),
            meanrich = mean(meanrich),
            precip = mean(precip),
            tmean = mean(tmean),
            tmax = mean(tmax),
            tmin = mean(tmin))
```

```{r tidy richness workspace, include=FALSE}
remove(mcLserp_grich)
remove(swan_grich)
remove(elk_grich)
remove(jasp_grich)
remove(ang_grich)
remove(McLAnn_grich)
remove(car_grich)
remove(ucsc_grich)
```

```{r guild rich graphs}
ggplot(all.grich.means, aes(x = year, y = meanrich, color = site)) +
  geom_line() +
  labs(title = "",
       x = "Year",
       y = "Richness") +
  facet_grid(~guild) 


guild.grich <- all.grich.means %>% filter(guild == "Exo" | guild == "Nat" | guild == "all")

grid.arrange(
  ggplot(guild.grich) +
  geom_point(aes(x = year, y = meanrich, color = guild)) +
  geom_line(aes(x = year, y = meanrich, color = guild)) +
    labs(title = "Richness of native and exotic species in 8 CA grasslands",
       x = "Year",
       y = "Richness") +
    facet_wrap(~site, nrow = 1))



exo.grich <- all.grich.means %>% filter(guild == "Exo")
nat.grich <- all.grich.means %>% filter(guild == "Nat")

exo.rich <- ggplot(nat.grich, aes(x = year, y = meanrich, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "Native species richness",
       x = "Year",
       y = "Richness") +
    facet_wrap(~site, nrow = 1)

nat.rich <- ggplot(exo.grich, aes(x = year, y = meanrich, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "Exotic species richness",
       x = "Year",
       y = "Richness") +
    facet_wrap(~site, nrow = 1)
```

## Relative richness of different guilds

```{r relative richness}
ang_grich <- angelo_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(plot, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAG = 0) %>% add_column(NAGr = 0) %>% add_column(EPG = 0) %>%
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "Angelo")

car_grich <- car_data %>%
  group_by(year, Plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, Plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(Plot, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(EPG = 0) %>% add_column(NPGr = 0) %>%
  add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "Carrizo")

elk_grich <- elkhorn_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(rep, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "Elkhorn")

jasp_grich <- jasper_data %>%
  group_by(year, uniqueID, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, uniqueID) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(uniqueID, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>% 
  add_column(NAGr = 0) %>%  add_column(EAF = 0) %>% add_column(EPF = 0) %>%
  add_column(EPG = 0) %>%  add_column(NPGr = 0) %>% add_column(NPS = 0) %>%
  add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "Jasper")

McLAnn_grich <- McLAnn_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich)  %>%
  select(plot, year, relrich, guild) %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "McL Ann")

mcLserp_grich <- mcLserp_data %>%
  group_by(year, plot, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich) %>%
  select(plot, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "McL Serp")

swan_grich <- swan_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich) %>%
  select(rep, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(EPG = 0) %>% 
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "Swanton")

ucsc_grich <- ucsc_data %>%
  group_by(year, rep, guild) %>%
  summarise(rich = n_distinct(species.name)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(totalrich = sum(rich),
         relrich = rich/totalrich) %>%
  select(rep, year, relrich, guild)  %>%
  spread(guild, relrich) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT),
         all = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT+EAF+EAG+EPF+EPG)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relrich") %>%
  group_by(year, guild) %>%
  summarise(meanrelrich = mean(relrich),
            serelrich = (std.error(relrich))) %>%
  mutate(site = "UCSC")

all.grelrich <- full_join(mcLserp_grich, swan_grich)
all.grelrich <- full_join(all.grelrich, elk_grich)
all.grelrich <- full_join(all.grelrich, jasp_grich)
all.grelrich <- full_join(all.grelrich, ang_grich)
all.grelrich <- full_join(all.grelrich, McLAnn_grich)
all.grelrich <- full_join(all.grelrich, car_grich)
all.grelrich <- full_join(all.grelrich, ucsc_grich)

all.grelrich <- left_join(all.grelrich, wtryr_weather, by = c("year", "site")) 

all.grelrich.means <- all.grelrich %>%
  group_by(site, guild, year) %>%
  summarise(serich = std.error(serelrich),
            meanrich = mean(meanrelrich),
            precip = mean(precip),
            tmean = mean(tmean),
            tmax = mean(tmax),
            tmin = mean(tmin))
```


```{r tidy richness workspace, include=FALSE}
remove(mcLserp_grich)
remove(swan_grich)
remove(elk_grich)
remove(jasp_grich)
remove(ang_grich)
remove(McLAnn_grich)
remove(car_grich)
remove(ucsc_grich)
```

```{r guild rich graphs}
ggplot(all.grelrich.means, aes(x = year, y = meanrich, color = site)) +
  geom_line() +
  labs(title = "",
       x = "Year",
       y = "Richness") +
  facet_grid(~guild) 

exo.relgrich <- all.grelrich.means %>% filter(guild == "Exo")
nat.relgrich <- all.grelrich.means %>% filter(guild == "Nat")


exo.relrich <- ggplot(nat.relgrich, aes(x = year, y = meanrich, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "Native species richness",
       x = "Year",
       y = "Richness") +
    facet_wrap(~site, nrow = 1)

nat.relrich <- ggplot(exo.relgrich, aes(x = year, y = meanrich, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "Exotic species richness",
       x = "Year",
       y = "Richness") +
    facet_wrap(~site, nrow = 1)

grid.arrange(exo.relrich, nat.relrich)

guild.grich <- all.grich.means %>% filter(guild == "Exo" | guild == "Nat")

grid.arrange(
  ggplot(guild.grich) +
  geom_point(aes(x = year, y = meanrich, color = guild)) +
  geom_line(aes(x = year, y = meanrich, color = guild)) +
    labs(title = "Richness of native and exotic species in 8 CA grasslands",
       x = "Year",
       y = "Relative Richness") +
    facet_wrap(~site, nrow = 1))
```

## Guild-richness analysis

This represents changes in overall richness and native vs. exotic richness over time. Because these sites all are measured in different ways, and I'm not particularly interested in measuring the differences between them, I will run separate analyses for each site. 

```{r Richness glmm}

glmer(Diversity ~ ppt.winter.s + (Year.s|Site), family = "poisson", data = DiversityFunc[DiversityFunc$FuncGroup == "Native Annuals",])

```



# Changes in guild-level cover over time

```{r converting species level cover to guild level cover}
ang_gcover <- angelo_data %>%
  group_by(year, plot, guild) %>%
  summarise(cover = sum(cover)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(plot, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAG = 0) %>% add_column(NAGr = 0) %>% add_column(EPG = 0) %>%
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "Angelo")

car_gcover <- car_data %>%
  group_by(year, Plot, guild) %>%
  summarise(cover = sum(Intercepts)) %>% ungroup() %>%
  group_by(year, Plot) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(Plot, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(EPG = 0) %>% add_column(NPGr = 0) %>%
  add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "Carrizo")

elk_gcover <- elkhorn_data %>%
  group_by(year, rep, guild) %>%
  summarise(cover = sum(intercepts)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(rep, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "Elkhorn")

jasp_gcover <- jasper_data %>%
  group_by(year, uniqueID, guild) %>%
  summarise(cover = sum(cover)) %>% ungroup() %>%
  group_by(year, uniqueID) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(uniqueID, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>% 
  add_column(NAGr = 0) %>%  add_column(EAF = 0) %>% add_column(EPF = 0) %>%
  add_column(EPG = 0) %>%  add_column(NPGr = 0) %>% add_column(NPS = 0) %>%
  add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "Jasper")

McLAnn_gcover <- McLAnn_data %>%
  filter(Cover > 0) %>%
  group_by(year, plot, guild) %>%
  summarise(cover = sum(Cover)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(plot, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "McL Ann")

mcLserp_gcover <- mcLserp_data %>%
  filter(Cover > 0) %>%
  group_by(year, plot, guild) %>%
  summarise(cover = sum(Cover)) %>% ungroup() %>%
  group_by(year, plot) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(plot, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAGr = 0) %>% add_column(NPGr = 0) %>% add_column(EPG = 0) %>% 
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "McL Serp")

swan_gcover <- swan_data %>%
  group_by(year, rep, guild) %>%
  summarise(cover = sum(intercepts)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(rep, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(EPG = 0) %>% 
  add_column(NPGr = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "Swanton")

ucsc_gcover <- ucsc_data %>%
  group_by(year, rep, guild) %>%
  summarise(cover = sum(intercepts)) %>% ungroup() %>%
  group_by(year, rep) %>%
  mutate(sumcov = sum(cover),
         relcov = cover/sumcov) %>% select(rep, year, relcov, guild)  %>%
  spread(guild, relcov) %>%
  replace(is.na(.), 0) %>%
  add_column(NAF = 0) %>% add_column(NAG = 0) %>% add_column(NPS = 0) %>% add_column(NPT = 0) %>%
  mutate(Nat = (NAF+NAG+NAGr+NPF+NPG+NPGr+NPS+NPT),
         Exo = (EAF+EAG+EPF+EPG),
         NatAnns = (NAF+NAG+NAGr),
         ExoAnns = (EAF+EAG),
         ExoPers = (EPF+EPG),
         NatPers = (NPF+NPG+NPGr+NPS+NPT)) %>%
  ungroup() %>% 
  gather('Nat', 'Exo', 'NatAnns', 'ExoAnns', 'ExoPers', 'NatPers', key = "guild", value = "relcov") %>%
  group_by(year, guild) %>%
  summarise(meancov = mean(relcov),
            secov = (std.error(relcov))) %>%
  mutate(site = "UCSC")

all.gcov <- full_join(mcLserp_gcover, swan_gcover)
all.gcov <- full_join(all.gcov, elk_gcover)
all.gcov <- full_join(all.gcov, jasp_gcover)
all.gcov <- full_join(all.gcov, ang_gcover)
all.gcov <- full_join(all.gcov, McLAnn_gcover)
all.gcov <- full_join(all.gcov, car_gcover)
all.gcov <- full_join(all.gcov, ucsc_gcover)

all_gcov <- left_join(all.gcov, wtryr_weather, by = c("year", "site")) 

all_gcov_means <- all_gcov %>%
  group_by(site, guild, year) %>%
  summarise(serelcov = std.error(meancov),
            meanrelcov = mean(meancov),
            precip = mean(precip),
            tmean = mean(tmean),
            tmax = mean(tmax),
            tmin = mean(tmin))
```

```{r tidy gcover workspace, include=FALSE}
remove(mcLserp_gcover, swan_gcover, elk_gcover, jasp_gcover, ang_gcover, McLAnn_gcover, car_gcover, ucsc_gcover)
```

```{r guild cover graphs}
ggplot(all_gcov_means, aes(x = year, y = meanrelcov, color = site)) +
  geom_line() +
  labs(title = "",
       x = "Year",
       y = "Relative cover") +
  facet_grid(~guild) 

exo.gcov <- all_gcov_means %>% filter(guild == "Exo")
nat.gcov <- all_gcov_means %>% filter(guild == "Nat")

grid.arrange(
  ggplot(nat.gcov, aes(x = year, y = meanrelcov, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "",
       x = "Year",
       y = "Relative cover"),
  ggplot(exo.gcov, aes(x = year, y = meanrelcov, color = site)) +
  geom_point() +
    geom_line() +
  labs(title = "",
       x = "Year",
       y = "Relative cover"))
```
